---
title: "CRPE Canopy Play File (Clean)"
author: "Anwesha Guha"
date: "5/13/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

# Clean Play File

## Cycle 3 Overview

This document contains analysis for items from the task tracker. The goal of this document is to consolidate the code that has already been done -- writing functions where necessary to streamline data analysis and cleaning processes.

First, read in data & load necessary libraries.

```{r libraries}
library(pacman)
p_load(tidyverse, here, rio, janitor, kableExtra, psych, DT)
```

```{r read in data}
data_dictionary <- import(here("data", "Final Canopy Dataset for DAC - Data dictionary.csv"), 
                          setclass = "tbl_df") %>% 
                  clean_names()
dac_22 <- import(here("data", "Final Canopy Dataset for DAC - Final Data 3_22.csv"),
                 setclass = "tbl_df") %>% 
            clean_names()
```


## Data in appropriate format

Pivot necessary columns from wide to long (will focus on practices/focus group -- assuming full data will be cleaned later)
```{r}
# Prep
# physical wellbeing was never selected as a core practice, so its data type is different from the others. converting it to the other data types for the sake of pivoting
dac_22$practices_physical_well_being <- as.character(dac_22$practices_physical_well_being)

#Pivot
df <- dac_22 %>%
  #pivot practices
  pivot_longer(cols = starts_with("practices_"),
               names_to = "practices",
               names_prefix = "practices_",
               values_to = "practice_frequency") %>% 
  #pivot focus groups
  #deselected free response question for analysis simplification -- should be revisited prior to any publication
  select(-focus_other_student_group_text) %>% 
  pivot_longer(cols = starts_with("focus_"),
               names_to = "focus",
               names_prefix = "focus_",
               values_to = "focus_frequency") %>% 
  #consolidating core practices to general practice_count column
  mutate(all_practice_count = case_when(
    practice_frequency == "0" ~ 0,
    TRUE ~ 1))
```

Simplify df to the ones I want to work with, focus area and practice counts.
```{r}
df2 <- df %>% 
  select(nces_id, 134:138, -practice_frequency)
```


Focus group descriptives.
```{r}
df3 <- df2 %>% 
  filter(focus_frequency >0, all_practice_count>0)
```

```{r}
focus_count <- df3 %>% 
  group_by(focus) %>% 
  summarise(n = n_distinct(nces_id))
```

```{r}
theme_set(theme_light())
focus_count %>% 
  ggplot(aes(x = reorder(focus, n), y = n)) +
  geom_col(fill = "navy", alpha = .8) +
  labs(x = "Focus Area Selected",
       y = "Practices Implemented",
       title = "Practices by Focus Area") +
  coord_flip()
```


Practice count descriptives.
```{r}
practice_count <- df3 %>% 
  group_by(practices) %>% 
  summarise(n = n_distinct(nces_id)) %>% 
  arrange(-n)
```

```{r}
practice_count %>% 
  datatable()
```
```{r}
head(practice_count$practices, 10)
```

schools focused on practices
unique number of practices 
practices implemented (mean practices per school)

bottom 5 practices for each focus 
% of focus area
reorder within level function -- sort practices within each facet
color overlap in a different color -- color by how many groups they are in

download transcend fonts

Total practice count by school
```{r}
practice_focus <- df3 %>% 
  group_by(practices, focus) %>% 
  summarise(n = sum(all_practice_count))
```

```{r}
practice_focus_plot <- practice_focus %>% 
  filter(practices %in% c(head(practice_count$practices, 10)))

library(colorblindr)
library(tidytext)

practice_focus %>% 
  group_by(focus) %>% 
  top_n(10) %>% 
  ungroup() %>% 
  mutate(perc = paste0(sprintf("%4.1f", n / sum(n) * 100), "%")) %>% 
  ggplot(aes(x = n, y = fct_reorder(practices, n), fill = focus)) + 
  geom_col(show.legend = FALSE, alpha = .8) +
#  geom_text(aes(label = perc)) +
  facet_wrap(~as.factor(focus), ncol = 3, scales = "free") +
  scale_fill_viridis_d() 
```

