---
title: 'CRPE Canopy Play: Cycle 2'
author: "Anwesha Guha"
date: "4/21/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)

library(here)
library(rio)
library(tidyverse)
library(janitor)
library(pracma)
library(kableExtra)
library(DT)
```

## Cycle 2: How do tagging patterns differ among schools designed for particular marginalized student groups?

### Import data

```{r data cleaning}
# read in data
data_dictionary <- import(here("data", "Final Canopy Dataset for DAC - Data dictionary.csv"), 
                          setclass = "tbl_df") %>% 
                  clean_names()
dac_22 <- import(here("data", "Final Canopy Dataset for DAC - Final Data 3_22.csv"),
                 setclass = "tbl_df") %>% 
            clean_names()
```

### Import relevant code from Janette's play file
*Core practices identified as "Your Most Important Practices" with up to 5 selected.*

```{r core practices}
# first check that I have all core practices accounted for and create variable name df
practices_var_names <- data_dictionary %>% 
  filter(row_number() %in% c(39:110))

# get sums
core_prac <- dac_22 %>% 
  as_tibble() %>% 
  select(nces_id, starts_with("practices_")) # 72 total practices

# pesky issues with one vector type
# convert integer vector to character to allow pivoting
core_prac$practices_physical_well_being <- as.character(core_prac$practices_physical_well_being)

# pivot 
core_prac <- core_prac %>% 
  pivot_longer(cols = starts_with("practices"),
               names_to = c("core_practice"),
               values_to = "practice_time") %>% 
  group_by(core_practice, practice_time) %>% 
  mutate(sum = n()) %>% 
  select(-nces_id) %>% 
  unique() %>% 
  pivot_wider(id_cols = "core_practice",
              names_from = "practice_time",
              values_from = "sum",
              names_repair = "minimal") %>% 
  select(c(1, 3, 2, 7, 6, 4, 5))

# code to double check the values
work_check <- dac_22 %>% 
  as_tibble() %>% 
  select(nces_id, starts_with("practices")) %>% 
  count(practices_students_leaders)

# values look good
# clean up NAs so they reflect 0 values
core_prac[is.na(core_prac)] <- 0

# final clean up for table
# prepare practices_var_names for joining
names(practices_var_names)[names(practices_var_names) == "variable_name"] <- "core_practice"

# prepare table df for joining and display
core_prac_freq <- core_prac %>% 
  select(-c(2,3)) %>% 
  left_join(practices_var_names, by = "core_practice") %>% 
  mutate(original_field_name_from_youssef_transcend = str_remove(original_field_name_from_youssef_transcend, "^practices_")) %>% 
  ungroup() %>% 
  select(6, 2, 3, 4, 5) %>%
  mutate(Total = rowSums(.[2:5]),
         Frequency = paste0(round((Total/sum(Total))*100, 2), "%"))

# rename var again
names(core_prac_freq)[names(core_prac_freq) == "original_field_name_from_youssef_transcend"] <- "Core Practice"
```

#### What are the frequencies of core practices?

```{r all frequency table}
datatable(core_prac_freq)
```

#### What are the frequencies of top 10 core practices?

```{r top 10 core}
top_10_core <- core_prac_freq %>% 
  arrange(Total) %>%
  slice_tail(n = 10)

datatable(top_10_core)
# double check accuracy
# max(core_prac_freq$Total)
# output is 50 -- bottom tail = 50
```

### Task: How do tagging patterns differ among schools designed for particular marginalized student groups?

#### Focus: BIPOC Students

```{r}
core_prac <- dac_22 %>% 
  as_tibble() %>% 
  filter(focus_bipoc == 1) %>% 
  select(nces_id, starts_with("practices_"))
```

```{r}
# pesky issues with one vector type
# convert integer vector to character to allow pivoting
core_prac$practices_physical_well_being <- as.character(core_prac$practices_physical_well_being)

# pivot 
core_prac <- core_prac %>% 
  pivot_longer(cols = starts_with("practices"),
               names_to = c("core_practice"),
               values_to = "practice_time") %>% 
  group_by(core_practice, practice_time) %>% 
  mutate(sum = n()) %>% 
  select(-nces_id) %>% 
  unique() %>% 
  pivot_wider(id_cols = "core_practice",
              names_from = "practice_time",
              values_from = "sum",
              names_repair = "minimal") %>% 
  select(c(1, 3, 2, 7, 6, 4, 5))

# code to double check the values
work_check <- dac_22 %>% 
  as_tibble() %>% 
  select(nces_id, starts_with("practices")) %>% 
  count(practices_students_leaders)

# values look good
# clean up NAs so they reflect 0 values
core_prac[is.na(core_prac)] <- 0

# final clean up for table
# prepare practices_var_names for joining
names(practices_var_names)[names(practices_var_names) == "variable_name"] <- "core_practice"

# prepare table df for joining and display
core_prac_freq <- core_prac %>% 
  select(-c(2,3)) %>% 
  left_join(practices_var_names, by = "core_practice") %>% 
  mutate(original_field_name_from_youssef_transcend = str_remove(original_field_name_from_youssef_transcend, "^practices_")) %>% 
  ungroup() %>% 
  select(6, 2, 3, 4, 5) %>%
  mutate(Total = rowSums(.[2:5]),
         Frequency = paste0(round((Total/sum(Total))*100, 2), "%"))

# rename var again
names(core_prac_freq)[names(core_prac_freq) == "original_field_name_from_youssef_transcend"] <- "Core Practice"

```


```{r all frequency table}
core_prac_freq %>% 
  arrange(-Total) %>% 
  datatable()
```

#### Focus: Economically Disadvantaged Students

```{r}
core_prac <- dac_22 %>% 
  as_tibble() %>% 
  filter(focus_economic_disadvantage == 1) %>% 
  select(nces_id, starts_with("practices_"))
```

```{r}
# pesky issues with one vector type
# convert integer vector to character to allow pivoting
core_prac$practices_physical_well_being <- as.character(core_prac$practices_physical_well_being)

# pivot 
core_prac <- core_prac %>% 
  pivot_longer(cols = starts_with("practices"),
               names_to = c("core_practice"),
               values_to = "practice_time") %>% 
  group_by(core_practice, practice_time) %>% 
  mutate(sum = n()) %>% 
  select(-nces_id) %>% 
  unique() %>% 
  pivot_wider(id_cols = "core_practice",
              names_from = "practice_time",
              values_from = "sum",
              names_repair = "minimal") %>% 
  select(c(1, 3, 2, 7, 6, 4, 5))

# code to double check the values
work_check <- dac_22 %>% 
  as_tibble() %>% 
  select(nces_id, starts_with("practices")) %>% 
  count(practices_students_leaders)

# values look good
# clean up NAs so they reflect 0 values
core_prac[is.na(core_prac)] <- 0

# final clean up for table
# prepare practices_var_names for joining
names(practices_var_names)[names(practices_var_names) == "variable_name"] <- "core_practice"

# prepare table df for joining and display
core_prac_freq <- core_prac %>% 
  select(-c(2,3)) %>% 
  left_join(practices_var_names, by = "core_practice") %>% 
  mutate(original_field_name_from_youssef_transcend = str_remove(original_field_name_from_youssef_transcend, "^practices_")) %>% 
  ungroup() %>% 
  select(6, 2, 3, 4, 5) %>%
  mutate(Total = rowSums(.[2:5]),
         Frequency = paste0(round((Total/sum(Total))*100, 2), "%"))

# rename var again
names(core_prac_freq)[names(core_prac_freq) == "original_field_name_from_youssef_transcend"] <- "Core Practice"

```


```{r all frequency table}
core_prac_freq %>% 
  arrange(-Total) %>% 
  datatable()
```

#### Focus: Emergent Bilinguals

```{r}
core_prac <- dac_22 %>% 
  as_tibble() %>% 
  filter(focus_emergent_bilinguals == 1) %>% 
  select(nces_id, starts_with("practices_"))
```

```{r}
# pesky issues with one vector type
# convert integer vector to character to allow pivoting
core_prac$practices_physical_well_being <- as.character(core_prac$practices_physical_well_being)

# pivot 
core_prac <- core_prac %>% 
  pivot_longer(cols = starts_with("practices"),
               names_to = c("core_practice"),
               values_to = "practice_time") %>% 
  group_by(core_practice, practice_time) %>% 
  mutate(sum = n()) %>% 
  select(-nces_id) %>% 
  unique() %>% 
  pivot_wider(id_cols = "core_practice",
              names_from = "practice_time",
              values_from = "sum",
              names_repair = "minimal") %>% 
  select(c(1, 3, 2, 7, 6, 4, 5))

# code to double check the values
work_check <- dac_22 %>% 
  as_tibble() %>% 
  select(nces_id, starts_with("practices")) %>% 
  count(practices_students_leaders)

# values look good
# clean up NAs so they reflect 0 values
core_prac[is.na(core_prac)] <- 0

# final clean up for table
# prepare practices_var_names for joining
names(practices_var_names)[names(practices_var_names) == "variable_name"] <- "core_practice"

# prepare table df for joining and display
core_prac_freq <- core_prac %>% 
  select(-c(2,3)) %>% 
  left_join(practices_var_names, by = "core_practice") %>% 
  mutate(original_field_name_from_youssef_transcend = str_remove(original_field_name_from_youssef_transcend, "^practices_")) %>% 
  ungroup() %>% 
  select(6, 2, 3, 4, 5) %>%
  mutate(Total = rowSums(.[2:5]),
         Frequency = paste0(round((Total/sum(Total))*100, 2), "%"))

# rename var again
names(core_prac_freq)[names(core_prac_freq) == "original_field_name_from_youssef_transcend"] <- "Core Practice"

```


```{r all frequency table}
core_prac_freq %>% 
  arrange(-Total) %>% 
  datatable()
```

#### Focus: Foster Students

```{r}
core_prac <- dac_22 %>% 
  as_tibble() %>% 
  filter(focus_foster == 1) %>% 
  select(nces_id, starts_with("practices_"))
```

```{r}
# pesky issues with one vector type
# convert integer vector to character to allow pivoting
core_prac$practices_physical_well_being <- as.character(core_prac$practices_physical_well_being)

# pivot 
core_prac <- core_prac %>% 
  pivot_longer(cols = starts_with("practices"),
               names_to = c("core_practice"),
               values_to = "practice_time") %>% 
  group_by(core_practice, practice_time) %>% 
  mutate(sum = n()) %>% 
  select(-nces_id) %>% 
  unique() %>% 
  pivot_wider(id_cols = "core_practice",
              names_from = "practice_time",
              values_from = "sum",
              names_repair = "minimal") %>% 
  select(c(1, 3, 2, 7, 6, 4, 5))

# code to double check the values
work_check <- dac_22 %>% 
  as_tibble() %>% 
  select(nces_id, starts_with("practices")) %>% 
  count(practices_students_leaders)

# values look good
# clean up NAs so they reflect 0 values
core_prac[is.na(core_prac)] <- 0

# final clean up for table
# prepare practices_var_names for joining
names(practices_var_names)[names(practices_var_names) == "variable_name"] <- "core_practice"

# prepare table df for joining and display
core_prac_freq <- core_prac %>% 
  select(-c(2,3)) %>% 
  left_join(practices_var_names, by = "core_practice") %>% 
  mutate(original_field_name_from_youssef_transcend = str_remove(original_field_name_from_youssef_transcend, "^practices_")) %>% 
  ungroup() %>% 
  select(6, 2, 3, 4, 5) %>%
  mutate(Total = rowSums(.[2:5]),
         Frequency = paste0(round((Total/sum(Total))*100, 2), "%"))

# rename var again
names(core_prac_freq)[names(core_prac_freq) == "original_field_name_from_youssef_transcend"] <- "Core Practice"

```


```{r all frequency table}
core_prac_freq %>% 
  arrange(-Total) %>% 
  datatable()
```

#### Focus: Homeless Students

```{r}
core_prac <- dac_22 %>% 
  as_tibble() %>% 
  filter(focus_homeless == 1) %>% 
  select(nces_id, starts_with("practices_"))
```

```{r}
# pesky issues with one vector type
# convert integer vector to character to allow pivoting
core_prac$practices_physical_well_being <- as.character(core_prac$practices_physical_well_being)

# pivot 
core_prac <- core_prac %>% 
  pivot_longer(cols = starts_with("practices"),
               names_to = c("core_practice"),
               values_to = "practice_time") %>% 
  group_by(core_practice, practice_time) %>% 
  mutate(sum = n()) %>% 
  select(-nces_id) %>% 
  unique() %>% 
  pivot_wider(id_cols = "core_practice",
              names_from = "practice_time",
              values_from = "sum",
              names_repair = "minimal") %>% 
  select(c(1, 3, 2, 7, 6, 4, 5))

# code to double check the values
work_check <- dac_22 %>% 
  as_tibble() %>% 
  select(nces_id, starts_with("practices")) %>% 
  count(practices_students_leaders)

# values look good
# clean up NAs so they reflect 0 values
core_prac[is.na(core_prac)] <- 0

# final clean up for table
# prepare practices_var_names for joining
names(practices_var_names)[names(practices_var_names) == "variable_name"] <- "core_practice"

# prepare table df for joining and display
core_prac_freq <- core_prac %>% 
  select(-c(2,3)) %>% 
  left_join(practices_var_names, by = "core_practice") %>% 
  mutate(original_field_name_from_youssef_transcend = str_remove(original_field_name_from_youssef_transcend, "^practices_")) %>% 
  ungroup() %>% 
  select(6, 2, 3, 4, 5) %>%
  mutate(Total = rowSums(.[2:5]),
         Frequency = paste0(round((Total/sum(Total))*100, 2), "%"))

# rename var again
names(core_prac_freq)[names(core_prac_freq) == "original_field_name_from_youssef_transcend"] <- "Core Practice"

```


```{r all frequency table}
core_prac_freq %>% 
  arrange(-Total) %>% 
  datatable()
```

#### Focus: Juvenile Justice

```{r}
core_prac <- dac_22 %>% 
  as_tibble() %>% 
  filter(focus_juvenile_justice == 1) %>% 
  select(nces_id, starts_with("practices_"))
```

```{r}
# pesky issues with one vector type
# convert integer vector to character to allow pivoting
core_prac$practices_physical_well_being <- as.character(core_prac$practices_physical_well_being)

# pivot 
core_prac <- core_prac %>% 
  pivot_longer(cols = starts_with("practices"),
               names_to = c("core_practice"),
               values_to = "practice_time") %>% 
  group_by(core_practice, practice_time) %>% 
  mutate(sum = n()) %>% 
  select(-nces_id) %>% 
  unique() %>% 
  pivot_wider(id_cols = "core_practice",
              names_from = "practice_time",
              values_from = "sum",
              names_repair = "minimal") %>% 
  select(c(1, 3, 2, 7, 6, 4, 5))

# code to double check the values
work_check <- dac_22 %>% 
  as_tibble() %>% 
  select(nces_id, starts_with("practices")) %>% 
  count(practices_students_leaders)

# values look good
# clean up NAs so they reflect 0 values
core_prac[is.na(core_prac)] <- 0

# final clean up for table
# prepare practices_var_names for joining
names(practices_var_names)[names(practices_var_names) == "variable_name"] <- "core_practice"

# prepare table df for joining and display
core_prac_freq <- core_prac %>% 
  select(-c(2,3)) %>% 
  left_join(practices_var_names, by = "core_practice") %>% 
  mutate(original_field_name_from_youssef_transcend = str_remove(original_field_name_from_youssef_transcend, "^practices_")) %>% 
  ungroup() %>% 
  select(6, 2, 3, 4, 5) %>%
  mutate(Total = rowSums(.[2:5]),
         Frequency = paste0(round((Total/sum(Total))*100, 2), "%"))

# rename var again
names(core_prac_freq)[names(core_prac_freq) == "original_field_name_from_youssef_transcend"] <- "Core Practice"

```


```{r all frequency table}
core_prac_freq %>% 
  arrange(-Total) %>% 
  datatable()
```

#### Focus: Newcomer Students (Immigrant and/or Refugees)

```{r}
core_prac <- dac_22 %>% 
  as_tibble() %>% 
  filter(focus_newcomer == 1) %>% 
  select(nces_id, starts_with("practices_"))
```

```{r}
# pesky issues with one vector type
# convert integer vector to character to allow pivoting
core_prac$practices_physical_well_being <- as.character(core_prac$practices_physical_well_being)

# pivot 
core_prac <- core_prac %>% 
  pivot_longer(cols = starts_with("practices"),
               names_to = c("core_practice"),
               values_to = "practice_time") %>% 
  group_by(core_practice, practice_time) %>% 
  mutate(sum = n()) %>% 
  select(-nces_id) %>% 
  unique() %>% 
  pivot_wider(id_cols = "core_practice",
              names_from = "practice_time",
              values_from = "sum",
              names_repair = "minimal") %>% 
  select(c(1, 3, 2, 7, 6, 4, 5))

# code to double check the values
work_check <- dac_22 %>% 
  as_tibble() %>% 
  select(nces_id, starts_with("practices")) %>% 
  count(practices_students_leaders)

# values look good
# clean up NAs so they reflect 0 values
core_prac[is.na(core_prac)] <- 0

# final clean up for table
# prepare practices_var_names for joining
names(practices_var_names)[names(practices_var_names) == "variable_name"] <- "core_practice"

# prepare table df for joining and display
core_prac_freq <- core_prac %>% 
  select(-c(2,3)) %>% 
  left_join(practices_var_names, by = "core_practice") %>% 
  mutate(original_field_name_from_youssef_transcend = str_remove(original_field_name_from_youssef_transcend, "^practices_")) %>% 
  ungroup() %>% 
  select(6, 2, 3, 4, 5) %>%
  mutate(Total = rowSums(.[2:5]),
         Frequency = paste0(round((Total/sum(Total))*100, 2), "%"))

# rename var again
names(core_prac_freq)[names(core_prac_freq) == "original_field_name_from_youssef_transcend"] <- "Core Practice"

```


```{r all frequency table}
core_prac_freq %>% 
  arrange(-Total) %>% 
  datatable()
```

#### Focus: Students with Learning Differences and Disabilities

```{r}
core_prac <- dac_22 %>% 
  as_tibble() %>% 
  filter(focus_swd == 1) %>% 
  select(nces_id, starts_with("practices_"))
```

```{r}
# pesky issues with one vector type
# convert integer vector to character to allow pivoting
core_prac$practices_physical_well_being <- as.character(core_prac$practices_physical_well_being)

# pivot 
core_prac <- core_prac %>% 
  pivot_longer(cols = starts_with("practices"),
               names_to = c("core_practice"),
               values_to = "practice_time") %>% 
  group_by(core_practice, practice_time) %>% 
  mutate(sum = n()) %>% 
  select(-nces_id) %>% 
  unique() %>% 
  pivot_wider(id_cols = "core_practice",
              names_from = "practice_time",
              values_from = "sum",
              names_repair = "minimal") %>% 
  select(c(1, 3, 2, 7, 6, 4, 5))

# code to double check the values
work_check <- dac_22 %>% 
  as_tibble() %>% 
  select(nces_id, starts_with("practices")) %>% 
  count(practices_students_leaders)

# values look good
# clean up NAs so they reflect 0 values
core_prac[is.na(core_prac)] <- 0

# final clean up for table
# prepare practices_var_names for joining
names(practices_var_names)[names(practices_var_names) == "variable_name"] <- "core_practice"

# prepare table df for joining and display
core_prac_freq <- core_prac %>% 
  select(-c(2,3)) %>% 
  left_join(practices_var_names, by = "core_practice") %>% 
  mutate(original_field_name_from_youssef_transcend = str_remove(original_field_name_from_youssef_transcend, "^practices_")) %>% 
  ungroup() %>% 
  select(6, 2, 3, 4, 5) %>%
  mutate(Total = rowSums(.[2:5]),
         Frequency = paste0(round((Total/sum(Total))*100, 2), "%"))

# rename var again
names(core_prac_freq)[names(core_prac_freq) == "original_field_name_from_youssef_transcend"] <- "Core Practice"

```


```{r all frequency table}
core_prac_freq %>% 
  arrange(-Total) %>% 
  datatable()
```
