---
title: "CRPE Canopy Play"
author: "Anwesha Guha"
date: "4/6/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(here)
library(rio)
library(tidyverse)
library(janitor)
library(pracma)
library(kableExtra)
```

```{r data cleaning}
# read in data
data_dictionary <- import(here("data", "Final Canopy Dataset for DAC - Data dictionary.csv"), 
                          setclass = "tbl_df") %>% 
                  clean_names()
dac_22 <- import(here("data", "Final Canopy Dataset for DAC - Final Data 3_22.csv"),
                 setclass = "tbl_df") %>% 
            clean_names()
```

What are the racial/ethnic and gender identities of Canopy school leaders?

First, gender.
```{r}
table(dac_22$confidential_leader_gender)
```

```{r}
dac_22$confidential_leader_gender[dac_22$confidential_leader_gender == "0"] <- "Not reported"
dac_22$confidential_leader_gender[dac_22$confidential_leader_gender == "Prefer not to say"] <- "Not reported"

dac_22 %>%
  ggplot(aes(confidential_leader_gender)) +
  geom_bar(aes(x = forcats::fct_infreq(confidential_leader_gender)), fill = "maroon4", alpha = .8) +
  theme_classic() +
  labs(title = "What are the gender identities of Canopy school leaders?",
       x = "Gender Identity",
       y = "Count")
```

```{r}
dac_22_race <- dac_22 %>% 
  pivot_longer(
    cols = starts_with("self_reported_race"),
    names_to = "self_reported_race",
    names_prefix = "self_reported_race_",
    values_to = "race_count"
  )
str(dac_22_race$race_count)
```

```{r}
race_grouped <- dac_22_race %>% 
  group_by(self_reported_race) %>% 
  summarise(race_count = sum(race_count, na.rm = TRUE))

race_grouped %>%
  ggplot(aes(self_reported_race, race_count)) +
  geom_col(aes(y = race_count, x = forcats::fct_infreq(self_reported_race)), fill = "maroon4", alpha = .8) +
  theme_classic() +
  labs(title = "What are the racial identities of Canopy school leaders?",
       subtitle = "*of those reporting race",
       x = "Racial Identity",
       y = "Count")
```

TYPE: How do tagging patterns differ among charter, district, and independent schools?
```{r}
dac_22$practices_physical_well_being <- as.character(dac_22$practices_physical_well_being) 
#kind of fascinating that this was the only factor that no schools ever chose in their top 5

dac_22_practices <- dac_22 %>% 
  pivot_longer(
    cols = starts_with("practices"),
    names_to = "practices",
    names_prefix = "practices_",
    values_to = "practice_count"
  )
```

Separating practice variables. Creating priority variable.
```{r}

dac_22_practices <- dac_22_practices %>% 
  mutate(practice_priority = practice_count) 

dac_22_practices$practice_priority[dac_22_practices$practice_priority == "0"] <- NA
dac_22_practices$practice_priority[dac_22_practices$practice_priority == "1"] <- NA
```

Converting other to numeric.
```{r}

dac_22_practices <- dac_22_practices %>% 
  mutate(practice_num = practice_count)

unique(dac_22_practices$practice_num)
dac_22_practices$practice_num[dac_22_practices$practice_num == "3-4 years"] <- 1
dac_22_practices$practice_num[dac_22_practices$practice_num == "5+ years"] <- 1
dac_22_practices$practice_num[dac_22_practices$practice_num == "1-2 years"] <- 1
dac_22_practices$practice_num[dac_22_practices$practice_num == "Less than a year"] <- 1

dac_22_practices$practice_num <- as.numeric(dac_22_practices$practice_num)
```

Group by practice & locale.
```{r}
str(dac_22_practices$practices)
practices_grouped_locale <- dac_22_practices %>% 
  group_by(practices, nces_locale) %>% 
  summarise(practice_tot = sum(practice_num, na.rm = TRUE))

practices_grouped_locale

practices_grouped <- dac_22_practices %>% 
  group_by(practices) %>% 
  summarise(practice_tot = sum(practice_num, na.rm = TRUE))

practices_grouped


```
```{r}
practices_grouped %>% 
  ggplot(aes(x = reorder(practices, practice_tot), practice_tot)) +
  geom_col() +
  coord_flip() +
  theme_classic()
  
```

```{r}
practices_grouped_locale %>% 
  subset(nces_locale == "Rural") %>% 
  arrange(desc(practice_tot)) %>% 
  head(15) %>% 
  ggplot(aes(x = reorder(practices, practice_tot), y = practice_tot)) +
  geom_col(fill = "maroon4", alpha = .8) +
  coord_flip() +
  theme_minimal() +
  labs(title = "Top Tags for Rural Schools", 
       x = "tags",
       y = "tag count")
```
```{r}
practices_grouped_locale %>% 
  subset(nces_locale == "Urban") %>% 
  arrange(desc(practice_tot)) %>% 
  head(15) %>% 
  ggplot(aes(x = reorder(practices, practice_tot), y = practice_tot)) +
  geom_col(fill = "maroon4", alpha = .8) +
  coord_flip() +
  theme_minimal() +
  labs(title = "Top Tags for Urban Schools",
       x = "tags",
       y = "tag count")
```

```{r}
practices_grouped_locale %>% 
  subset(nces_locale == "Suburban") %>% 
  arrange(desc(practice_tot)) %>% 
  head(15) %>% 
  ggplot(aes(x = reorder(practices, practice_tot), y = practice_tot)) +
  geom_col(fill = "maroon4", alpha = .8) +
  coord_flip() +
  theme_minimal() +
  labs(title = "Top Tags for Suburban Schools", 
       x = "tags",
       y = "tag count")
```

How do tagging patterns differ among schools designed for particular marginalized student groups?

