---
title: "Frequency Differences - BIPOC"
author: "Anwesha Guha"
date: "6/15/2022"
output: 
  html_document:
    toc: true
    toc_float: true
    theme: flatly
    highlight: tango
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Overview of tasks

Tags that have...

* the biggest frequency differences (meaning, % of schools selecting the tag) between 2 categories
* tags with the smallest frequency differences
* biggest odds ratios for selecting tags

For the following ... 

* Schools led by a leader of color vs. not
* Schools w/ >75% BIPOC leadership teams vs. <75%
* Schools w/ >50% BIPOC leadership teams vs. <50%

# Odds Ratios 

## Setup for analysis

```{r}
library(pacman)
p_load(rio, here, psych, tidyverse, DT, scales, extrafont)
source(here("scripts", "branding.R"))
load(here("data", "complete_canopy_2022.RData"))
```

## Schools led by leader of color or not

### Setup

BIPOC leaders are all non-White school leaders, including those self-identifying as another race. Only one NA value was removed. This school (Aspen Academy) may have a multiple leader model.

*Note:* 

```{r}
leader_bipoc_tags <- clean_data_full %>% 
  mutate(leader_bipoc = ifelse(confidential_leader_race == "White", "white", "bipoc")) %>% 
  select(school_id, leader_bipoc, starts_with("practices")) %>% 
  subset(!is.na(leader_bipoc)) %>% 
  pivot_longer(cols = starts_with("practices"),
               names_to = "tag",
               values_to = "n") %>% 
  select(-school_id) %>% 
  group_by(leader_bipoc, tag) %>% 
  mutate(n = sum(n, na.rm = TRUE)) %>% 
  ungroup() %>% 
  unique() %>% 
#percents for each practice by leader_bipoc
  group_by(leader_bipoc) %>% 
  mutate(total_tags_leader = sum(n),
         pct_school = n/160, 
         pct_total_tags = n/total_tags_leader) #schools in each leader per tag / total tages per leader - ignore for odds
```

### Odds ratio

```{r leader odds}
leader_odds <- leader_bipoc_tags %>% 
  select(leader_bipoc, tag, pct_school) %>% 
  pivot_wider(names_from = "leader_bipoc",
              values_from = "pct_school",
              names_glue = "{.value}_{leader_bipoc}") %>% 
  janitor::clean_names() %>% 
  mutate(odds_white = pct_school_white / (1 - pct_school_white),
         odds_bipoc = pct_school_bipoc / (1 - pct_school_bipoc),
         bipoc_white_ratio = odds_bipoc / odds_white)
```

```{r bipoc v white odd viz, fig.dim = c(9, 9)}
bipoc_white_odds <- leader_odds %>% 
  select(tag, bipoc_white_ratio) %>% 
  slice_max(bipoc_white_ratio, n = 15, with_ties = FALSE)

#odd ratio labels
bipoc_white_odds_lab <- data.frame(x = c(bipoc_white_odds$tag),
                           y = c(bipoc_white_odds$bipoc_white_ratio + .05),
                           lab = c(paste0(round(bipoc_white_odds$bipoc_white_ratio, 1), "x")))

#viz
bipoc_white_odds_viz <- bipoc_white_odds %>% 
  ggplot(aes(reorder(tag, bipoc_white_ratio), bipoc_white_ratio)) +
  geom_bar(stat = "identity", position = "dodge", fill = "#1A4C81") +
  coord_flip() +
  geom_text(aes(x = bipoc_white_odds_lab$x, y = bipoc_white_odds_lab$y, label = bipoc_white_odds_lab$lab), 
            color = "#EF464B", size = 5, family = "Bebas Neue") +
  scale_color_manual(values = paste(transcend_cols)) +
  scale_x_tag() +
  theme_transcend +
  theme(axis.text.x = element_blank(),
        axis.ticks.x = element_blank(),
        axis.text = element_text(size = 7)) +
  labs(x = "",
       y = "",
       title = "How Much More Likely BIPOC School \nLeaders were to Select a Tag",
       subtitle = "using raw odds ratio by highest difference")
  
bipoc_white_odds_viz
```

## Leadership Team: Schools w/ >75% BIPOC leadership teams vs. <75%

### Setup
```{r}
table(clean_data_full$confidential_leadership_team_diversity)
```


```{r}
bipoc75_tags <- clean_data_full %>% 
  mutate(bipoc_leadership = ifelse(confidential_leadership_team_diversity == "75 - 100% people of color", "Mostly BIPOC", "Not Mostly BIPOC")) %>% 
  select(school_id, bipoc_leadership, starts_with("practices")) %>% 
  subset(!is.na(bipoc_leadership)) %>% 
  pivot_longer(cols = starts_with("practices"),
               names_to = "tag",
               values_to = "n") %>% 
  select(-school_id) %>% 
  group_by(bipoc_leadership, tag) %>% 
  mutate(n = sum(n, na.rm = TRUE)) %>% 
  ungroup() %>% 
  unique() %>% 
#percents for each practice by bipoc_leadership
  group_by(bipoc_leadership) %>% 
  mutate(total_tags_bipoc = sum(n),
         pct_school = n/161, 
         pct_total_tags = n/total_tags_bipoc) #mostly bipoc leadership schools per tag / total tags per bipoc leadership - ignore for odds
```

### Odds ratio

```{r bipoc75 leadership odds}
bipoc75_odds <- bipoc75_tags %>% 
  select(bipoc_leadership, tag, pct_school) %>% 
  pivot_wider(names_from = "bipoc_leadership",
              values_from = "pct_school",
              names_glue = "{.value}_{bipoc_leadership}") %>% 
  janitor::clean_names() %>% 
  mutate(odds_mostly_bipoc = pct_school_mostly_bipoc / (1 - pct_school_mostly_bipoc),
         odds_notmostly_bipoc = pct_school_not_mostly_bipoc / (1 - pct_school_not_mostly_bipoc),
         most_notmost_bipoc_ratio = odds_mostly_bipoc / odds_notmostly_bipoc)
```

```{r bipoc75 v less odd viz, fig.dim = c(9, 9)}
top_bipoc75_odds <- bipoc75_odds %>% 
  select(tag, most_notmost_bipoc_ratio) %>% 
  slice_max(most_notmost_bipoc_ratio, n = 15, with_ties = FALSE)

#odd ratio labels
top_bipoc75_odds_lab <- data.frame(x = c(top_bipoc75_odds$tag),
                           y = c(top_bipoc75_odds$most_notmost_bipoc_ratio + .05),
                           lab = c(paste0(round(top_bipoc75_odds$most_notmost_bipoc_ratio, 1), "x")))

#viz
bipoc75_odds_viz <- top_bipoc75_odds %>% 
  ggplot(aes(reorder(tag, most_notmost_bipoc_ratio), most_notmost_bipoc_ratio)) +
  geom_bar(stat = "identity", position = "dodge", fill = "#1A4C81") +
  coord_flip() +
  geom_text(aes(x = top_bipoc75_odds_lab$x, y = top_bipoc75_odds_lab$y, label = top_bipoc75_odds_lab$lab), 
            color = "#EF464B", size = 5, family = "Bebas Neue") +
  scale_color_manual(values = paste(transcend_cols)) +
  scale_x_tag() +
  theme_transcend +
  theme(axis.text.x = element_blank(),
        axis.ticks.x = element_blank(),
        axis.text = element_text(size = 7)) +
  labs(x = "",
       y = "",
       title = "How Much More Likely 75% BIPOC Leadership \nSchools were to Select a Tag",
       subtitle = "using raw odds ratio by highest difference")
  
bipoc75_odds_viz
```

## Leadership Team: Schools w/ >50% BIPOC leadership teams vs. <50%

### Setup
```{r}
bipoc50_tags <- clean_data_full %>% 
  mutate(bipoc_leadership = case_when(
    confidential_leadership_team_diversity == "50 - 74% people of color" ~ "Majority BIPOC",
    confidential_leadership_team_diversity == "75 - 100% people of color" ~ "Majority BIPOC",
    TRUE ~ "Not Majority BIPOC"
  )) %>% 
  select(school_id, bipoc_leadership, starts_with("practices")) %>%
  subset(!is.na(bipoc_leadership)) %>% 
  pivot_longer(cols = starts_with("practices"),
               names_to = "tag",
               values_to = "n") %>% 
  select(-school_id) %>% 
  group_by(bipoc_leadership, tag) %>% 
  mutate(n = sum(n, na.rm = TRUE)) %>% 
  ungroup() %>% 
  unique() %>% 
#percents for each practice by bipoc_leadership
  group_by(bipoc_leadership) %>% 
  mutate(total_tags_bipoc = sum(n),
         pct_school = n/161, 
         pct_total_tags = n/total_tags_bipoc) #mostly bipoc leadership schools per tag / total tags per bipoc leadership - ignore for odds
```

### Odds ratio

```{r bipoc50 leadership odds}
bipoc50_odds <- bipoc50_tags %>% 
  select(bipoc_leadership, tag, pct_school) %>% 
  pivot_wider(names_from = "bipoc_leadership",
              values_from = "pct_school",
              names_glue = "{.value}_{bipoc_leadership}") %>% 
  janitor::clean_names() %>% 
  mutate(odds_majority_bipoc = pct_school_majority_bipoc / (1 - pct_school_majority_bipoc),
         odds_notmajority_bipoc = pct_school_not_majority_bipoc / (1 - pct_school_not_majority_bipoc),
         majority_notmajority_bipoc_ratio = odds_majority_bipoc / odds_notmajority_bipoc)
```

```{r bipoc50 v less odd viz, fig.dim = c(9, 9)}
top_bipoc50_odds <- bipoc50_odds %>% 
  select(tag, majority_notmajority_bipoc_ratio) %>% 
  slice_max(majority_notmajority_bipoc_ratio, n = 15, with_ties = FALSE)

#odd ratio labels
top_bipoc50_odds_lab <- data.frame(x = c(top_bipoc50_odds$tag),
                           y = c(top_bipoc50_odds$majority_notmajority_bipoc_ratio + .05),
                           lab = c(paste0(round(top_bipoc50_odds$majority_notmajority_bipoc_ratio, 1), "x")))

#viz
bipoc50_odds_viz <- top_bipoc50_odds %>% 
  ggplot(aes(reorder(tag, majority_notmajority_bipoc_ratio), majority_notmajority_bipoc_ratio)) +
  geom_bar(stat = "identity", position = "dodge", fill = "#1A4C81") +
  coord_flip() +
  geom_text(aes(x = top_bipoc50_odds_lab$x, y = top_bipoc50_odds_lab$y, label = top_bipoc50_odds_lab$lab), 
            color = "#EF464B", size = 5, family = "Bebas Neue") +
  scale_color_manual(values = paste(transcend_cols)) +
  scale_x_tag() +
  theme_transcend +
  theme(axis.text.x = element_blank(),
        axis.ticks.x = element_blank(),
        axis.text = element_text(size = 7)) +
  labs(x = "",
       y = "",
       title = "How Much More Likely 50% BIPOC Leadership \nSchools were to Select a Tag",
       subtitle = "using raw odds ratio by highest difference")
  
bipoc50_odds_viz
```


