---
title: "Janette's Data Sandbox"
author: "Janette Avelar"
date: "3/25/2022"
output: 
  html_document:
    toc: true
    toc_float: true
    theme: flatly
    highlight: tango
    code_folding: hide
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
```

```{r library, include = FALSE}
library(here)
library(rio)
library(tidyverse)
library(janitor)
library(pracma)
library(kableExtra)
library(DT)
```

```{r data cleaning}
# read in data
data_dictionary <- import(here("data", "Final Canopy Dataset for DAC - Data dictionary.csv"), 
                          setclass = "tbl_df") %>% 
                  clean_names()
dac_22 <- import(here("data", "Final Canopy Dataset for DAC - Final Data 3_22.csv"),
                 setclass = "tbl_df") %>% 
            clean_names()
```
   
# Cycle 1 (4.8.22)

Guiding questions/tasks & format:  
1. What are the frequencies of core practices?  
   - table  
2. How long have schools been implementing each core practice?  
   - table or stacked bar chart  
3. What are the frequencies of all practices?  
   - table  
   
Task 1: Frequencies

*Core practices identified as "Your Most Important Practices" with up to 5 selected.*

```{r core practices}
# first check that I have all core practices accounted for and create variable name df
practices_var_names <- data_dictionary %>% 
  filter(row_number() %in% c(39:110))

# get sums
core_prac <- dac_22 %>% 
  as_tibble() %>% 
  select(nces_id, starts_with("practices_")) # 72 total practices

# pesky issues with one vector type
# convert integer vector to character to allow pivoting
core_prac$practices_physical_well_being <- as.character(core_prac$practices_physical_well_being)

# pivot 
core_prac <- core_prac %>% 
  pivot_longer(cols = starts_with("practices"),
               names_to = c("core_practice"),
               values_to = "practice_time") %>% 
  group_by(core_practice, practice_time) %>% 
  mutate(sum = n()) %>% 
  select(-nces_id) %>% 
  unique() %>% 
  pivot_wider(id_cols = "core_practice",
              names_from = "practice_time",
              values_from = "sum",
              names_repair = "minimal") %>% 
  select(c(1, 3, 2, 7, 6, 4, 5))

# code to double check the values
work_check <- dac_22 %>% 
  as_tibble() %>% 
  select(nces_id, starts_with("practices")) %>% 
  count(practices_students_leaders)

# values look good
# clean up NAs so they reflect 0 values
core_prac[is.na(core_prac)] <- 0

# final clean up for table
# prepare practices_var_names for joining
names(practices_var_names)[names(practices_var_names) == "variable_name"] <- "core_practice"

# prepare table df for joining and display
core_prac_freq <- core_prac %>% 
  select(-c(2,3)) %>% 
  left_join(practices_var_names, by = "core_practice") %>% 
  mutate(original_field_name_from_youssef_transcend = str_remove(original_field_name_from_youssef_transcend, "^practices_")) %>% 
  ungroup() %>% 
  select(6, 2, 3, 4, 5) %>%
  mutate(Total = rowSums(.[2:5]),
         Frequency = paste0(round((Total/sum(Total))*100, 2), "%"))

# rename var again
names(core_prac_freq)[names(core_prac_freq) == "original_field_name_from_youssef_transcend"] <- "Core Practice"
```

## What are the frequencies of core practices?

```{r all frequency table}
datatable(core_prac_freq)
```

## What are the frequencies of top 10 core practices?

```{r top 10 core}
top_10_core <- core_prac_freq %>% 
  arrange(Total) %>%
  slice_tail(n = 10)

datatable(top_10_core)
# double check accuracy
# max(core_prac_freq$Total)
# output is 50 -- bottom tail = 50
```

## How long have schools been implementing each core practice?

```{r stack chart for implementation times}
#prep data to tidy format
core_imp_time <- top_10_core %>% 
  select(-c(6)) %>% #drop last col
  pivot_longer(cols = c(2:5), #pivot implementation times to column
               names_to = "implementation_time",
               values_to = "N")

#manually refactor implementation_time column
core_imp_time$implementation_time <- as.factor(core_imp_time$implementation_time)
# order "Less than a year" first
core_imp_time$implementation_time <- relevel(core_imp_time$implementation_time, ref = "Less than a year")

#plot
core_imp_time %>% 
  ggplot(aes(reorder(`Core Practice`, -N), N, fill = implementation_time)) +
  geom_bar(position = "stack", stat = "identity") +
  scale_fill_manual(values = c("#76b81b","#ff0000", "#00ced1", "#8a2be2"),
                    name = "Implementation Time") +
  theme(axis.text.x = element_text(angle = 45, hjust = 0.95, vjust = 0.95),
        axis.text.y = element_text(angle = 90)) +
  scale_x_discrete(name = "") +
  scale_y_discrete(name = "Number of schools \n (2021 and 2022)")
```

## What are the frequencies of all practices?

```{r freq all practices}
all_prac_freq <- core_prac %>% 
  select(c(1, 3:7)) %>% # drop unseen practices due to display logic 
  group_by(core_practice) %>% 
  mutate(Total = sum(`1`, `Less than a year`, `1-2 years`, `3-4 years`, `5+ years`)) %>% # core + chosen practice
  ungroup() %>% 
  mutate(Frequency = paste0(round((Total/sum(Total))*100, 2), "%")) %>% #calculate freq
  select(-c(2:6)) %>%  # drop core freqs
  left_join(practices_var_names, by = "core_practice") %>% #join labels
  mutate(original_field_name_from_youssef_transcend = str_remove(original_field_name_from_youssef_transcend, "^practices_")) %>% #drop naming convention for clarity in table
  select(4, 2, 3) %>% #drop redundant row
  rename("Practice Selected" = original_field_name_from_youssef_transcend) #rename col 1
```

```{r all practices table}
datatable(all_prac_freq)
```

## What are the frequencies for the top 10 of all practices?

```{r top 10 all practices}
#filter top 10 practices
top_10_prac <- all_prac_freq %>% 
  slice_max(order_by = Frequency, n = 10)


## double check
# max(all_prac_freq$Frequency) #130 matches top in table of 130
```

```{r top 10 of all practices freq}
datatable(top_10_prac)
```

# Cycle 2 (4.22.22)

To Do:  
* Create new locale variable from NCES and self-reported data  
* Collapse school type variables  
* Less important: Pivot practices (double check sums are correct)  
* Collapse leader race variable (will need if/else for bi/multiracial variable)  
* Fill meaningful data (yes/no/NA)

## Data Cleaning  

### Locale

Created a new variable `locale` to indicate whether a school is urban, rural, or suburban. This variable merges NCES and self-reported data, with NCES data taking precedence in instances where we had information from both sources. There is 1 instance of `NA` where neither NCES or self-reported data were recorded.

*Status:* In progress - 10 schools have multi-coded self-reported data (specified below). I need to know how to code these in order to move forward with factoring the variable for easier analysis.

```{r locale variable}
# !diagnostics off
# Note a '1' indicates a yes
clean_dac_22 <- dac_22 #new set

#relabel empty NCES data as NA
clean_dac_22$nces_locale[clean_dac_22$nces_locale == ""] <- NA

#merge NCES and self-reported data
clean_dac_22 <- clean_dac_22 %>% 
  unite("locale", c(nces_locale, self_reported_locale_rural, self_reported_locale_suburban, self_reported_locale_urban), na.rm = TRUE) %>% 
  ungroup()
#relabeling new locale data
clean_dac_22$locale[clean_dac_22$locale == ""] <- NA #empties/all NA
clean_dac_22$locale[clean_dac_22$locale == "_1_0_0"] <- "Rural" #self-reported rural
clean_dac_22$locale[clean_dac_22$locale == "_0_1_0"] <- "Suburban" #self-reported suburban
clean_dac_22$locale[clean_dac_22$locale == "_0_0_1"] <- "Urban" #self-reported urban
clean_dac_22$locale[clean_dac_22$locale == "Urban_0_0_1"] <- "Urban" #NCES overriding self-report

unique(clean_dac_22$locale) #double check all labels are accounted for
#the following are unaccounted for labels that need to be specified and the particular instances where they arose
#issues coming from self-reporting multiple locale types by row
View(clean_dac_22$locale)
View(dac_22[156:157,]) #_1_1_1 [all 3]
View(dac_22[123,]) 
View(dac_22[140,])
View(dac_22[114,]) #_0_1_1 [suburban and urban]
View(dac_22[132:133,]) 
View(dac_22[144,]) #_1_0_1 [rural and urban]
View(dac_22[131,])
View(dac_22[147,])

#rows to clean up specified below:
which((clean_dac_22$locale == "_1_1_1") |
        (clean_dac_22$locale == "_0_1_1") |
      (clean_dac_22$locale == "_1_0_1"))

#once these final rows are cleaned up we can factor this var
sum(is.na(clean_dac_22$locale))
```

### School Type

School type was previously divided into 10 columns representing the following:  
* district  
* charter  
* independent  
* sws  
* virtual  
* microschool  
* homeschool  
* hub  
* other  
* text description for `other`

Values representing yes and no have been condensed into a single column named `school_type` specifying one of the values from the list above.

*Status:* In progress - a few instances of `other` and another item from list chosen. I need to know what to do in instances of multi-coding in order to proceed. Also need to know if `other` text should be its own value, or if a value of `other` with an additional column for text is preferable.

```{r collapse school type}
#district c131
#charter c132
#independent c133
#sws c134
#virtual c135
#microschool c136
#homeschool c137
#hub c138
#other c139
#col 140 is text description


# View(summarize(count(clean_dac_22[,131:139])))
# match("school_descriptor_district", names(clean_dac_22))
# unique(clean_dac_22$school_descriptor_other_text)
# which(clean_dac_22$school_descriptor_other_text == "& Hybrid")
# which(clean_dac_22$school_descriptor_other_text == "RI State Public School")
# which(clean_dac_22$school_descriptor_other_text == "Public School that serves deaf, blind, and deafblind students from 0 - 22")
View(dac_22[152,])
View(dac_22[157:158,])

#work smaller dataset to ensure code works properly
#will also use this to ensure changes to clean dataset match up
schl_check <- clean_dac_22 %>% 
  select(school_name, c(131:139)) %>% 
  group_by(school_name) %>% 
  mutate(overlap = n()) #output incorrect
#need to think about dually-ID'd school types.

# unique(schl_check$overlap)
View(schl_check[152,])
```

### School Leader Racial Identity

School leaders filling out the survey were asked to report how they identify racially. This variable was previously coded across 9 columns, with the following values:  
* AIAN  
* Asian  
* Black  
* Hispanic  
* NHPI  
* White  
* Self describe option  
* Text for `self describe`  
* No Response (which I assume is prefer not to respond)

Values were coded as either `yes` or `no` which has been collapsed into a single column `confidential_leader_race` with a value filled in according to the list above. An additional value, `Biracial` has been added for leaders who identified as more than one race (there were no instances with more than 2 races reported).  

The new values include:  
* White  
* Biracial  
* Hispanic  
* Asian  
* No Response  
* NA (1 missing observation)  
* Xicana Indigena (1 observation)  
* Middle Eastern (1 observation)  
* AIAN  

Note that NHPI was dropped because there were no coded NHPI cases. Self-reported text was used, but I can change this to "Other" or to any other value desired if needed.

*Status:* Complete -- variable is a factor with previously listed values.

```{r leader race collapsed, eval = FALSE}
#cols 141-148
#self describe text col 10

#look at uniquely identified self-report race data
unique(clean_dac_22$confidential_leader_race_self_describe_text)
#which observations?
which(clean_dac_22$confidential_leader_race_self_describe_text  == "xicana indigena") #row 114
which(clean_dac_22$confidential_leader_race_self_describe_text  == "Middle Eastern") #row 121
which(clean_dac_22$confidential_leader_race_self_describe_text  == "-99") #row 161

clean_dac_22$confidential_leader_race_aian[clean_dac_22$confidential_leader_race_aian == "American Indian or Alaska Native"] <- 1
clean_dac_22$confidential_leader_race_aian <- as.integer(clean_dac_22$confidential_leader_race_aian)
#how many bi/multiracial participants? do we need a var?
race_check <- clean_dac_22 %>% 
  select(school_name, c(141:148, 10)) %>% 
  mutate(Total = rowSums(.[2:8]))
#check unique values
unique(race_check$Total)
#at most, participants chose 2 racial identities
#new category will be "biracial"

#collapsing race in clean dataset
clean_dac_22 <- as.data.frame(clean_dac_22) %>% 
  unite("confidential_leader_race",
        c(confidential_leader_race_aian, confidential_leader_race_asian,
        confidential_leader_race_black, confidential_leader_race_hispanic,
        confidential_leader_race_nhpi, confidential_leader_race_white,
        confidential_leader_gender_self_describe, confidential_leader_race_self_describe_text,
        confidential_leader_race_no_response), na.rm = TRUE)
#recode column so it makes sense  
clean_dac_22<- clean_dac_22 %>% 
  mutate(confidential_leader_race = recode(clean_dac_22$confidential_leader_race,
         "1_0_0_0_0_0_0_0_0" = "AIAN",
         "0_1_0_0_0_0_0_0_0" = "Asian",
         "0_0_1_0_0_0_0_0_0" = "Black",
         #"American Indian or Alaska Native_0_0_0_0_1_0_0_0" = "Biracial", #AIAN + White (2)
         #"American Indian or Alaska Native_0_1_0_0_0_0_0_0" = "Biracial", #AIAN + Black (1)
         #"American Indian or Alaska Native_0_0_1_0_0_0_0_0" = "Biracial", #AIAN + Hispanic (1)
         "1_0_0_0_0_1_0_0_0" = "Biracial", #AIAN + White (2)
         "0_0_0_1_0_1_0_0_0" = "Biracial", #Hispanic + White (2)
         "0_0_1_1_0_0_0_0_0" = "Biracial", #Hispanic + Black (1)
         "0_0_1_0_0_1_0_0_0" = "Biracial", #Black + White (1)
         "1_0_0_1_0_0_0_0_0" = "Biracial", #AIAN + Hispanic
         "1_0_1_0_0_0_0_0_0" = "Biracial", #AIAN + Black (1)
         "0_0_0_1_0_0_0_0_0" = "Hispanic",
         #"XXX" = "NHPI", there were 0 coded NHPI
         "0_0_0_0_0_1_0_0_0" = "White",
         "0_0_0_0_0_1_-99_-99_0" = "White", #not sure about -99?
         "0_0_0_0_0_0_0_xicana indigena_0" = "Xicana Indigena", #self describe text (1)
         "0_0_0_0_0_0_0_Middle Eastern_0" = "Middle Eastern", #self describe text (1)
         "0_0_0_0_0_0_0_0_1" = "No Response", #(2)
         "0_0_0_0_0_0_0_0_0" = "NA"), #(1)
  confidential_leader_race = factor(confidential_leader_race,
                                           levels = c("AIAN", "Asian", "Black", "Biracial", "Hispanic", "White", "Xicana Indigena", "Middle Eastern", "No Response", "NA")))
```

### Private School NCES Data

The NCES data merged with our Canopy dataset includes counts for each grade level within private schools. Each variable (listed below) has been recoded so that any values are converted to "Yes", "No", "Not Reported", or "NA" as applicable. (Note NAs were used when cells were completely empty. NA is **not** the same as Not Reported.)  

Variables of interest:  
* `nces_grades_pre_k`  
* `nces_grades_kindergarten`  
* `nces_grades_1st`  
* `nces_grades_2nd`  
* `nces_grades_3rd`  
* `nces_grades_4th`  
* `nces_grades_5th`  
* `nces_grades_6th`  
* `nces_grades_7th`  
* `nces_grades_8th`  
* `nces_grades_9th`  
* `nces_grades_10th`  
* `nces_grades_11th`  
* `nces_grades_12th`  
* `nces_grades_13th`  

I did not create a new variable designating grade levels offered (e.g., elementary, middle, high, mixed elementary/jr high, mixed jr/sr high, etc.) but can do so if needed.

Can also recode so NA is the same as Not Reported if needed.

*Status:* Complete.

```{r private grade recode, eval = FALSE}
#manually refactor NAs
#refactor() not working for empty cells marked as ""
clean_dac_22$nces_grades_pre_k[clean_dac_22$nces_grades_pre_k == ""] <- NA
clean_dac_22$nces_grades_kindergarten[clean_dac_22$nces_grades_kindergarten == ""] <- NA
clean_dac_22$nces_grades_1st[clean_dac_22$nces_grades_1st == ""] <- NA
clean_dac_22$nces_grades_2nd[clean_dac_22$nces_grades_2nd == ""] <- NA
clean_dac_22$nces_grades_3rd[clean_dac_22$nces_grades_3rd == ""] <- NA
clean_dac_22$nces_grades_4th[clean_dac_22$nces_grades_4th == ""] <- NA
clean_dac_22$nces_grades_5th[clean_dac_22$nces_grades_5th == ""] <- NA
clean_dac_22$nces_grades_6th[clean_dac_22$nces_grades_6th == ""] <- NA
clean_dac_22$nces_grades_7th[clean_dac_22$nces_grades_7th == ""] <- NA
clean_dac_22$nces_grades_8th[clean_dac_22$nces_grades_8th == ""] <- NA
clean_dac_22$nces_grades_9th[clean_dac_22$nces_grades_9th == ""] <- NA
clean_dac_22$nces_grades_10th[clean_dac_22$nces_grades_10th == ""] <- NA
clean_dac_22$nces_grades_11th[clean_dac_22$nces_grades_11th == ""] <- NA
clean_dac_22$nces_grades_12th[clean_dac_22$nces_grades_12th == ""] <- NA
clean_dac_22$nces_grades_13th[clean_dac_22$nces_grades_13th == ""] <- NA
#recode all numeric values to "Yes" and factor
clean_dac_22 <- clean_dac_22 %>% 
  mutate(nces_grades_pre_k = recode(temp_clean_dac_22$nces_grades_pre_k,
                                    "10" = "Yes",
                                    "33" = "Yes",
                                    "8" = "Yes",
                                    "16" = "Yes",
                                    "13" = "Yes"),
         nces_grades_kindergarten = recode(temp_clean_dac_22$nces_grades_kindergarten,
                                           "4" = "Yes",
                                           "28" = "Yes",
                                           "15" = "Yes",
                                           "3" = "Yes",
                                           "8" = "Yes",
                                           "17" = "Yes"),
         nces_grades_1st = recode(temp_clean_dac_22$nces_grades_1st,
                                  "4" = "Yes",
                                  "1" = "Yes",
                                  "33" = "Yes",
                                  "13" = "Yes",
                                  "3" = "Yes",
                                  "11" = "Yes",
                                  "15" = "Yes",
                                  "14" = "Yes"),
         nces_grades_2nd = recode(temp_clean_dac_22$nces_grades_2nd,
                                  "6" = "Yes",
                                  "1" = "Yes",
                                  "41" = "Yes",
                                  "13" = "Yes",
                                  "4" = "Yes",
                                  "8" = "Yes"),
         nces_grades_3rd = recode(temp_clean_dac_22$nces_grades_3rd,
                                  "5" = "Yes",
                                  "47" = "Yes",
                                  "8" = "Yes",
                                  "13" = "Yes",
                                  "21" = "Yes"),
         nces_grades_4th = recode(temp_clean_dac_22$nces_grades_4th,
                                  "5" = "Yes",
                                  "1" = "Yes",
                                  "38" = "Yes",
                                  "18" = "Yes",
                                  "7" = "Yes",
                                  "12" = "Yes",
                                  "19" = "Yes",
                                  "21" = "Yes"),
         nces_grades_5th = recode(temp_clean_dac_22$nces_grades_5th,
                                  "7" = "Yes",
                                  "5" = "Yes",
                                  "52" = "Yes",
                                  "16" = "Yes",
                                  "10" = "Yes",
                                  "20" = "Yes"),
         nces_grades_6th = recode(temp_clean_dac_22$nces_grades_6th,
                                  "3" = "Yes",
                                  "45" = "Yes",
                                  "15" = "Yes",
                                  "8" = "Yes",
                                  "10" = "Yes",
                                  "16" = "Yes",
                                  "22" = "Yes"),
         nces_grades_7th = recode(temp_clean_dac_22$nces_grades_7th,
                                  "4" = "Yes",
                                  "2" = "Yes",
                                  "46" = "Yes",
                                  "11" = "Yes",
                                  "9" = "Yes",
                                  "6" = "Yes",
                                  "13" = "Yes",
                                  "22" = "Yes"),
         nces_grades_8th = recode(temp_clean_dac_22$nces_grades_8th,
                                  "6" = "Yes",
                                  "1" = "Yes",
                                  "42" = "Yes",
                                  "10" = "Yes",
                                  "7" = "Yes",
                                  "11" = "Yes",
                                  "16" = "Yes"),
         nces_grades_9th = recode(temp_clean_dac_22$nces_grades_9th,
                                  "4" = "Yes",
                                  "35" = "Yes",
                                  "2" = "Yes",
                                  "3" = "Yes",
                                  "21" = "Yes"),
         nces_grades_10th = recode(temp_clean_dac_22$nces_grades_10th,
                                   "9" = "Yes",
                                   "44" = "Yes",
                                   "2" = "Yes",
                                   "8" = "Yes"),
         nces_grades_11th = recode(temp_clean_dac_22$nces_grades_11th,
                                   "5" = "Yes",
                                   "17" = "Yes",
                                   "7" = "Yes",
                                   "6" = "Yes"),
         nces_grades_12th = recode(temp_clean_dac_22$nces_grades_12th,
                                   "4" = "Yes",
                                   "1" = "Yes",
                                   "5" = "Yes"),
         nces_grades_pre_k = factor(nces_grades_pre_k),
         nces_grades_kindergarten = factor(nces_grades_kindergarten),
         nces_grades_1st = factor(nces_grades_1st),
         nces_grades_2nd = factor(nces_grades_2nd),
         nces_grades_3rd = factor(nces_grades_3rd),
         nces_grades_4th = factor(nces_grades_4th),
         nces_grades_5th = factor(nces_grades_5th),
         nces_grades_6th = factor(nces_grades_6th),
         nces_grades_7th = factor(nces_grades_7th),
         nces_grades_8th = factor(nces_grades_8th),
         nces_grades_9th = factor(nces_grades_9th),
         nces_grades_10th = factor(nces_grades_10th),
         nces_grades_11th = factor(nces_grades_11th),
         nces_grades_12th = factor(nces_grades_12th),
         nces_grades_13th = factor(nces_grades_13th))
```


