---
title: "Janette's Data Sandbox"
author: "Janette Avelar"
date: "3/25/2022"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r library}
library(here)
library(rio)
library(tidyverse)
library(janitor)
library(pracma)
library(kableExtra)
```

```{r data cleaning}
# read in data
data_dictionary <- import(here("data", "Final Canopy Dataset for DAC - Data dictionary.csv"), 
                          setclass = "tbl_df") %>% 
                  clean_names()
dac_22 <- import(here("data", "Final Canopy Dataset for DAC - Final Data 3_22.csv"),
                 setclass = "tbl_df") %>% 
            clean_names()
```

Guiding questions/tasks & format:
1. What are the frequencies of core practices?
   - table
2. How long have schools been implementing each core practice?
   - table or stacked bar chart
3. ~if time~ What's the breakdown of schools in the data by:
   - locale (urban, suburban, rural)
   - level (elementary, middle, high)
   - school type (charter, district, independent)
4. ~if time as #1~ What are the frequencies of all practices?
   - table
   
# Round 1 (4.8.22)
## Task 1: What are the frequencies of core practices?

*Core practices identified as "Your Most Important Practices" with up to 5 selected.*

Create new cols with mutate that sum
- where ==1
- where ==Less than 1 year
etc.
- also where ==0

```{r core practices}
# first check that I have all core practices accounted for and create variable name df
practices_var_names <- data_dictionary %>% 
  filter(row_number() %in% c(39:110))

# get sums
core_prac <- dac_22 %>% 
  as_tibble() %>% 
  select(nces_id, starts_with("practices_")) # 72 total practices

# pesky issues with one vector type
# convert integer vector to character to allow pivoting
core_prac$practices_physical_well_being <- as.character(core_prac$practices_physical_well_being)
typeof(core_prac$practices_physical_well_being)

# pivot 
core_prac <- core_prac %>% 
  pivot_longer(cols = starts_with("practices"),
               names_to = c("core_practice"),
               values_to = "practice_time") %>% 
  group_by(core_practice, practice_time) %>% 
  mutate(sum = n()) %>% 
  select(-nces_id) %>% 
  unique() %>% 
  pivot_wider(id_cols = "core_practice",
              names_from = "practice_time",
              values_from = "sum",
              names_repair = "minimal") %>% 
  select(c(1, 3, 2, 7, 6, 4, 5))

# code to double check the values
work_check <- dac_22 %>% 
  as_tibble() %>% 
  select(nces_id, starts_with("practices")) %>% 
  count(practices_students_leaders)

# values look good
# clean up NAs so they reflect 0 values
core_prac[is.na(core_prac)] <- 0

# final clean up for table
# prepare practices_var_names for joining
names(practices_var_names)[names(practices_var_names) == "variable_name"] <- "core_practice"

# prepare table df for joining and display
core_prac_freq <- core_prac %>% 
  select(-c(2,3)) %>% 
  left_join(practices_var_names, by = "core_practice") %>% 
  mutate(original_field_name_from_youssef_transcend = str_remove(original_field_name_from_youssef_transcend, "^practices_")) %>% 
  ungroup() %>% 
  select(6, 2, 3, 4, 5) %>%
  mutate(Total = rowSums(.[2:5]))

# rename var again
names(core_prac_freq)[names(core_prac_freq) == "original_field_name_from_youssef_transcend"] <- "Core Practice"
```

### What are the frequencies of core practices?

```{r core frequency table}
kable(core_prac_freq)
```

*Note to self to clean that table up--it's ugly!*

### Stacked Bar Chart for Implementation

```{r stack chart for implementation times}
core_imp_time <- core_prac_freq %>% 
  ggplot()
```

