---
title: "Janette's Data Sandbox"
author: "Janette Avelar"
date: "3/25/2022"
output: 
  html_document:
    toc: true
    toc_float: true
    theme: flatly
    highlight: tango
    code_folding: hide
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
```

```{r library, include = FALSE}
library(here)
library(rio)
library(tidyverse)
library(janitor)
library(pracma)
library(kableExtra)
library(DT)
```

```{r data cleaning}
# read in data
data_dictionary <- import(here("data", "Final Canopy Dataset for DAC - Data dictionary.csv"), 
                          setclass = "tbl_df") %>% 
                  clean_names()
dac_22 <- import(here("data", "Final Canopy Dataset for DAC - Final Data 3_22.csv"),
                 setclass = "tbl_df") %>% 
            clean_names()
```
   
# Cycle 1 (4.8.22)

Guiding questions/tasks & format:  
1. What are the frequencies of core practices?  
   - table  
2. How long have schools been implementing each core practice?  
   - table or stacked bar chart  
3. What are the frequencies of all practices?  
   - table  
   
Task 1: Frequencies

*Core practices identified as "Your Most Important Practices" with up to 5 selected.*

```{r core practices}
# first check that I have all core practices accounted for and create variable name df
practices_var_names <- data_dictionary %>% 
  filter(row_number() %in% c(39:110))

# get sums
core_prac <- dac_22 %>% 
  as_tibble() %>% 
  select(nces_id, starts_with("practices_")) # 72 total practices

# pesky issues with one vector type
# convert integer vector to character to allow pivoting
core_prac$practices_physical_well_being <- as.character(core_prac$practices_physical_well_being)

# pivot 
core_prac <- core_prac %>% 
  pivot_longer(cols = starts_with("practices"),
               names_to = c("core_practice"),
               values_to = "practice_time") %>% 
  group_by(core_practice, practice_time) %>% 
  mutate(sum = n()) %>% 
  select(-nces_id) %>% 
  unique() %>% 
  pivot_wider(id_cols = "core_practice",
              names_from = "practice_time",
              values_from = "sum",
              names_repair = "minimal") %>% 
  select(c(1, 3, 2, 7, 6, 4, 5))

# code to double check the values
work_check <- dac_22 %>% 
  as_tibble() %>% 
  select(nces_id, starts_with("practices")) %>% 
  count(practices_students_leaders)

# values look good
# clean up NAs so they reflect 0 values
core_prac[is.na(core_prac)] <- 0

# final clean up for table
# prepare practices_var_names for joining
names(practices_var_names)[names(practices_var_names) == "variable_name"] <- "core_practice"

# prepare table df for joining and display
core_prac_freq <- core_prac %>% 
  select(-c(2,3)) %>% 
  left_join(practices_var_names, by = "core_practice") %>% 
  mutate(original_field_name_from_youssef_transcend = str_remove(original_field_name_from_youssef_transcend, "^practices_")) %>% 
  ungroup() %>% 
  select(6, 2, 3, 4, 5) %>%
  mutate(Total = rowSums(.[2:5]),
         Frequency = paste0(round((Total/sum(Total))*100, 2), "%"))

# rename var again
names(core_prac_freq)[names(core_prac_freq) == "original_field_name_from_youssef_transcend"] <- "Core Practice"
```

## What are the frequencies of core practices?

```{r all frequency table}
datatable(core_prac_freq)
```

## What are the frequencies of top 10 core practices?

```{r top 10 core}
top_10_core <- core_prac_freq %>% 
  arrange(Total) %>%
  slice_tail(n = 10)

datatable(top_10_core)
# double check accuracy
# max(core_prac_freq$Total)
# output is 50 -- bottom tail = 50
```

## How long have schools been implementing each core practice?

```{r stack chart for implementation times}
#prep data to tidy format
core_imp_time <- top_10_core %>% 
  select(-c(6)) %>% #drop last col
  pivot_longer(cols = c(2:5), #pivot implementation times to column
               names_to = "implementation_time",
               values_to = "N")

#manually refactor implementation_time column
core_imp_time$implementation_time <- as.factor(core_imp_time$implementation_time)
# order "Less than a year" first
core_imp_time$implementation_time <- relevel(core_imp_time$implementation_time, ref = "Less than a year")

#plot
core_imp_time %>% 
  ggplot(aes(reorder(`Core Practice`, -N), N, fill = implementation_time)) +
  geom_bar(position = "stack", stat = "identity") +
  scale_fill_manual(values = c("#76b81b","#ff0000", "#00ced1", "#8a2be2"),
                    name = "Implementation Time") +
  theme(axis.text.x = element_text(angle = 45, hjust = 0.95, vjust = 0.95),
        axis.text.y = element_text(angle = 90)) +
  scale_x_discrete(name = "") +
  scale_y_discrete(name = "Number of schools \n (2021 and 2022)")
```

## What are the frequencies of all practices?

```{r freq all practices}
all_prac_freq <- core_prac %>% 
  select(c(1, 3:7)) %>% # drop unseen practices due to display logic 
  group_by(core_practice) %>% 
  mutate(Total = sum(`1`, `Less than a year`, `1-2 years`, `3-4 years`, `5+ years`)) %>% # core + chosen practice
  ungroup() %>% 
  mutate(Frequency = paste0(round((Total/sum(Total))*100, 2), "%")) %>% #calculate freq
  select(-c(2:6)) %>%  # drop core freqs
  left_join(practices_var_names, by = "core_practice") %>% #join labels
  mutate(original_field_name_from_youssef_transcend = str_remove(original_field_name_from_youssef_transcend, "^practices_")) %>% #drop naming convention for clarity in table
  select(4, 2, 3) %>% #drop redundant row
  rename("Practice Selected" = original_field_name_from_youssef_transcend) #rename col 1
```

```{r all practices table}
datatable(all_prac_freq)
```

## What are the frequencies for the top 10 of all practices?

```{r top 10 all practices}
#filter top 10 practices
top_10_prac <- all_prac_freq %>% 
  slice_max(order_by = Frequency, n = 10)


## double check
# max(all_prac_freq$Frequency) #130 matches top in table of 130
```

```{r top 10 of all practices freq}
datatable(top_10_prac)
```

# Cycle 2 (4.22.22)

To Do:  
* Create new locale variable from NCES and self-reported data  
* Collapse school type variables  
* Less important: Pivot practices (double check sums are correct)  
* Collapse leader race variable (will need if/else for bi/multiracial variable)  
* Fill meaningful data (yes/no/NA)

## Data Cleaning  
```{r locale variable}
# !diagnostics off
# Note a '1' indicates a yes
clean_dac_22 <- dac_22 #new set

#relabel empty NCES data as NA
clean_dac_22$nces_locale[clean_dac_22$nces_locale == ""] <- NA

#merge NCES and self-reported data
clean_dac_22 <- clean_dac_22 %>% 
  unite("locale", c(nces_locale, self_reported_locale_rural, self_reported_locale_suburban, self_reported_locale_urban), na.rm = TRUE) %>% 
  ungroup()
#relabeling new locale data
clean_dac_22$locale[clean_dac_22$locale == ""] <- NA #empties/all NA
clean_dac_22$locale[clean_dac_22$locale == "_1_0_0"] <- "Rural" #self-reported rural
clean_dac_22$locale[clean_dac_22$locale == "_0_1_0"] <- "Suburban" #self-reported suburban
clean_dac_22$locale[clean_dac_22$locale == "_0_0_1"] <- "Urban" #self-reported urban
clean_dac_22$locale[clean_dac_22$locale == "Urban_0_0_1"] <- "Urban" #NCES overriding self-report

unique(clean_dac_22$locale) #double check all labels are accounted for
#the following are unaccounted for labels that need to be specified and the particular instances where they arose
#issues coming from self-reporting multiple locale types
View(clean_dac_22$locale)
View(dac_22[156:157,]) #_1_1_1 [all 3]
View(dac_22[123,]) 
View(dac_22[140,])
View(dac_22[114,]) #_0_1_1 [suburban and urban]
View(dac_22[132:133,]) 
View(dac_22[144,]) #_1_0_1 [rural and urban]
View(dac_22[131,])
View(dac_22[147,])

which((clean_dac_22$locale == "_1_1_1") |
        (clean_dac_22$locale == "_0_1_1") |
      (clean_dac_22$locale == "_1_0_1"))
```

```{r collapse school type}
#district c131
#charter c132
#independent c133
#sws c134
#virtual c135
#microschool c136
#homeschool c137
#hub c138
#other c139
#col 140 is text description
View(summarize(count(clean_dac_22[,131:139])))
match("school_descriptor_district", names(clean_dac_22))
unique(clean_dac_22$school_descriptor_other_text)
which(clean_dac_22$school_descriptor_other_text == "& Hybrid")
which(clean_dac_22$school_descriptor_other_text == "RI State Public School")
which(clean_dac_22$school_descriptor_other_text == "Public School that serves deaf, blind, and deafblind students from 0 - 22")
View(dac_22[152,])
View(dac_22[157:158,])

schl_check <- clean_dac_22 %>% 
  select(school_name, c(131:139)) %>% 
  group_by(school_name) %>% 
  mutate(overlap = n()) #output incorrect
#need to think about dually-ID'd school types.

unique(schl_check$overlap)
View(schl_check[152,])
```

