---
title: "Level Analysis"
author: "Janette Avelar"
date: "5/23/2022"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
```

```{r library}
library(here)
library(rio)
library(psych)
library(tidyverse)
library(ggplot2)
library(DT)
library(scales)
#source("scripts/branding.R", local = knitr::knit_global())
load(here("data", "complete_canopy_2022.RData"))
```

# Cycle 4 (5.24.22) [in preparation for Brown Bag]

Guiding questions/tasks & format:  
`LEVEL`: How do tagging patterns differ among elementary, middle, and high schools?  
* table of biggest differences between categories (elem vs. middle; middle vs. high; elem vs. high)  
* odds ratio  
* visualization  

Analysis Notes:  
* elementary schools = 43  
* middle schools = 59  
* high schools = 90  
* elementary AND middle = 31  
* middle AND high = 29  
* elementary AND middle AND high = 16  
* NONE elementary middle or high = 19  
* total schools = 157


*Focus on odds ratio plot for levels by Tuesday--other plots by this weekend*

```{r data set up and clean, include = FALSE}
level_tags <- clean_data_full %>% 
  select(school_id, grades_elementary, grades_middle, grades_high, starts_with("practices")) %>% 
  pivot_longer(cols = starts_with("practices"),
               names_to = "tag",
               values_to = "n")
which(level_tags$grades_high==0 & level_tags$grades_middle==0 & level_tags$grades_elementary==0)
View(data.frame(colnames(levels)))
```

