---
title: "Level Analysis"
author: "Janette Avelar"
date: "5/23/2022"
output: 
  html_document:
    toc: true
    toc_float: true
    theme: flatly
    highlight: tango
    code_folding: hide
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
```

```{r library}
library(here)
library(rio)
library(psych)
library(tidyverse)
library(ggplot2)
library(DT)
library(scales)
source(here("scripts", "branding.R")) #, local = knitr::knit_global())
load(here("data", "complete_canopy_2022.RData"))
```

# Cycle 4 (5.24.22) [in preparation for Brown Bag]

Guiding questions/tasks & format:  
`LEVEL`: How do tagging patterns differ among elementary, middle, and high schools?  
* table of biggest differences between categories (elem vs. middle; middle vs. high; elem vs. high)  
* odds ratio  
* visualization  

## Analysis Notes:  
* elementary schools = 50  
* middle schools = 67  
* high schools = 93  
* elementary AND middle = 38  
* middle AND high = 32  
* elementary AND middle AND high = 19  
* NONE elementary middle or high = 11  
    * CodeRVA Regional High School   
    * Empower Community High School  
    * Future Public School  
    * Purdue Polytechnic High School  
    * Rooted School  
    * Living School  
    * Achieve Community Charter School  
    * Southeast Area Technical High School (Sea-Tech)  
    * Sequoyah School  
    * Vimenti School  
    * Foundation School Of Autism - San Antonio  
* 10 `NA` values
* total schools = 161 - 10 `NA` = 151


*Focus on odds ratio plot for levels by Tuesday--other plots by this weekend*

```{r data set up and clean, include = FALSE}
level_tags <- clean_data_full %>% 
  select(school_id, grades_elementary, grades_middle, grades_high, starts_with("practices")) %>% 
  filter(!c(is.na(grades_elementary) & is.na(grades_middle) & is.na(grades_high))) %>% 
  mutate(grades_elementary = ifelse(grades_elementary==1, "elementary", "none"),
         grades_middle = ifelse(grades_middle==1, "middle", "none"),
         grades_high = ifelse(grades_high==1, "high", "none")) %>% 
  pivot_longer(cols = starts_with("grades"),
               names_to = "og_level",
               values_to = "level") %>% 
  select(-og_level) %>% 
  filter(!level=="none") %>% 
  pivot_longer(cols = starts_with("practices"),
               names_to = "tag",
               values_to = "n") %>% 
  select(-school_id) %>% 
  group_by(level, tag) %>% 
  mutate(n = sum(n, na.rm = TRUE)) %>% 
  ungroup() %>% 
  unique() %>% 
#percents for each practice by level
  group_by(level) %>% 
  mutate(total_tags_level = sum(n),
         pct_school = n/151, #only 151 schools in the analysis
         pct_total_tags = n/total_tags_level) #schools in each locale per tag / total tags in locale - ignore for odds
```

## Probability and Odds Ratios

In this final analysis, I calculated probability for each practice and converted to odds ratio in order to plot how much more or less likely a given locale was to select each tag.

```{r level odds}
level_odds <- level_tags %>% 
  select(level, tag, pct_school) %>% 
  pivot_wider(names_from = "level",
              values_from = "pct_school",
              names_glue = "{.value}_{level}") %>% 
  janitor::clean_names() %>% 
  mutate(odds_elem = pct_school_elementary / (1 - pct_school_elementary),
         odds_mid = pct_school_middle / (1 - pct_school_middle),
         odds_high = pct_school_high / (1 - pct_school_high),
         elem_mid_ratio = odds_elem / odds_mid,
         mid_high_ratio = odds_mid / odds_high,
         high_elem_ratio = odds_high / odds_elem)
```

### Elementary vs. Middle (odds ratio)

```{r elem v mid odd viz, fig.dim = c(9, 9)}
elem_mid_odds <- level_odds %>% 
  select(tag, elem_mid_ratio) %>% 
  slice_max(elem_mid_ratio, n = 15, with_ties = FALSE)
#odd ratio labels
elem_mid_odds_lab <- data.frame(x = c(elem_mid_odds$tag),
                           y = c(elem_mid_odds$elem_mid_ratio + .05),
                           lab = c(paste0(round(elem_mid_odds$elem_mid_ratio, 1), "x")))
#viz
elem_mid_odds_viz <- elem_mid_odds %>% 
  ggplot(aes(reorder(tag, -elem_mid_ratio), elem_mid_ratio)) +
  geom_bar(stat = "identity", position = "dodge", fill = "#1A4C81") +
  coord_flip() +
  geom_text(aes(x = elem_mid_odds_lab$x, y = elem_mid_odds_lab$y, label = elem_mid_odds_lab$lab), 
            color = "#EF464B", size = 7, family = "Bebas Neue") +
  scale_color_manual(values = paste(transcend_cols)) +
  scale_x_tag() +
  theme_transcend +
  theme(axis.text.x = element_blank(),
        axis.ticks.x = element_blank()) +
  labs(x = "",
       y = "",
       title = "How Much More Likely Elementary Schools were to \nSelect a Tag",
       subtitle = "compared to middle schools using raw odds ratio")
  
elem_mid_odds_viz
```

### Middle vs. High (odds ratio)

```{r mid v high odd viz, fig.dim = c(9, 9)}
mid_high_odds <- level_odds %>% 
  select(tag, mid_high_ratio) %>% 
  slice_max(mid_high_ratio, n = 15, with_ties = FALSE)
#odd ratio labels
mid_high_odds_lab <- data.frame(x = c(mid_high_odds$tag),
                           y = c(mid_high_odds$mid_high_ratio + .05),
                           lab = c(paste0(round(mid_high_odds$mid_high_ratio, 1), "x")))
#viz
mid_high_odds_viz <- mid_high_odds %>% 
  ggplot(aes(reorder(tag, -mid_high_ratio), mid_high_ratio)) +
  geom_bar(stat = "identity", position = "dodge", fill = "#1A4C81") +
  coord_flip() +
  geom_text(aes(x = mid_high_odds_lab$x, y = mid_high_odds_lab$y, label = mid_high_odds_lab$lab), 
            color = "#EF464B", size = 7, family = "Bebas Neue") +
  scale_color_manual(values = paste(transcend_cols)) +
  scale_x_tag() +
  theme_transcend +
  theme(axis.text.x = element_blank(),
        axis.ticks.x = element_blank()) +
  labs(x = "",
       y = "",
       title = "How Much More Likely Middle Schools were to \nSelect a Tag",
       subtitle = "compared to high schools using raw odds ratio")
  
mid_high_odds_viz
```

### High vs. Elementary (odds ratio)

```{r high v elem odd viz, fig.dim = c(9, 9)}
high_elem_odds <- level_odds %>% 
  select(tag, high_elem_ratio) %>% 
  slice_max(high_elem_ratio, n = 15, with_ties = FALSE)
#odd ratio labels
high_elem_odds_lab <- data.frame(x = c(high_elem_odds$tag),
                           y = c(high_elem_odds$high_elem_ratio + .35),
                           lab = c(paste0(round(high_elem_odds$high_elem_ratio, 1), "x")))
#viz
high_elem_odds_viz <- high_elem_odds %>% 
  ggplot(aes(reorder(tag, -high_elem_ratio), high_elem_ratio)) +
  geom_bar(stat = "identity", position = "dodge", fill = "#1A4C81") +
  coord_flip() +
  geom_text(aes(x = high_elem_odds_lab$x, y = high_elem_odds_lab$y, label = high_elem_odds_lab$lab), 
            color = "#EF464B", size = 7, family = "Bebas Neue") +
  scale_color_manual(values = paste(transcend_cols)) +
  scale_x_tag() +
  theme_transcend +
  theme(axis.text.x = element_blank(),
        axis.ticks.x = element_blank()) +
  labs(x = "",
       y = "",
       title = "How Much More Likely High Schools were to \nSelect a Tag",
       subtitle = "compared to elementary schools using raw odds ratio")
  
high_elem_odds_viz
```

Save plots:
```{r, eval = FALSE}
#odds charts
ggsave("elem-mid_odds_chart_raw.png", plot = elem_mid_odds_viz, path = here("outputs", "level analysis"),
       width = 12, height = 8, units = "in")
ggsave("mid-high_odds_chart_raw.png", plot = mid_high_odds_viz, path = here("outputs", "level analysis"),
       width = 12, height = 8, units = "in")
ggsave("high-elem_odds_chart_raw.png", plot = high_elem_odds_viz, path = here("outputs", "level analysis"),
       width = 12, height = 8, units = "in")
```

# Cycle 4 (6.3.22)

Notes:  
* Elementary Schools = 50  
* Middle Schools = 67  
* High Schools = 93  
* Per previous discussion, categories were not kept mutually exclusive  

All visualizations reflect within-category differences, but a table of across-category differences is also included.

## Level Differences Table (Across)

```{r prep diffs}
level_diffs <- level_tags %>% 
  select(level, tag, n, pct_school) %>% 
  pivot_wider(names_from = "level",
              values_from = c("n", "pct_school")) %>% 
  mutate(elem_mid_diff = abs(round(100*(pct_school_elementary - pct_school_middle), 2)),
         mid_high_diff = abs(round(100*(pct_school_middle - pct_school_high), 2)),
         high_elem_diff = abs(round(100*(pct_school_high - pct_school_elementary), 2)))
#tbl
level_diffs_tbl <- level_diffs %>% 
  select(tag, elem_mid_diff, mid_high_diff, high_elem_diff, pct_school_elementary, pct_school_middle, pct_school_high, n_elementary, n_middle, n_high) %>% 
  mutate(pct_school_elementary = paste0(round(100*pct_school_elementary, 2), "%"),
         pct_school_middle = paste0(round(100*pct_school_middle, 2), "%"),
         pct_school_high = paste0(round(100*pct_school_high, 2), "%"),
         elem_mid_diff = paste0(elem_mid_diff, "%"),
         mid_high_diff = paste0(mid_high_diff, "%"),
         high_elem_diff = paste0(high_elem_diff, "%")) %>% 
  left_join(tag_labels, by = "tag") %>% 
  select(c(label, elem_mid_diff, mid_high_diff, high_elem_diff, pct_school_elementary, pct_school_middle, pct_school_high, n_elementary, n_middle, n_high)) %>% 
  rename(`Tag` = label,
         `Elementary vs. Middle Difference` = elem_mid_diff,
         `Middle vs. High Difference` = mid_high_diff,
         `High vs. Elementary Difference` = high_elem_diff,
         `Pct. Tags (Elementary)` = pct_school_elementary,
         `Pct. Tags (Middle)` = pct_school_middle,
         `Pct. Tags (High)` = pct_school_high,
         `N Tags (Elementary)` = n_elementary,
         `N Tags (Middle)` = n_middle,
         `N Tags (High)` = n_high)
#table
datatable(level_diffs_tbl)
```

## Level Differences Table (Within)

```{r level diff tbl within}
#Get number of schools in each category
counts <- clean_data_full %>% 
  select(school_id, grades_elementary, grades_middle, grades_high, starts_with("practices")) %>% 
  filter(!c(is.na(grades_elementary) & is.na(grades_middle) & is.na(grades_high))) %>% 
  mutate(grades_elementary = ifelse(grades_elementary==1, "elementary", "none"),
         grades_middle = ifelse(grades_middle==1, "middle", "none"),
         grades_high = ifelse(grades_high==1, "high", "none")) %>% 
  pivot_longer(cols = starts_with("grades"),
               names_to = "og_level",
               values_to = "level") %>% 
  select(-og_level) %>% 
  filter(!level=="none") %>% 
  select(level) %>% 
  summarize(n_sch_elem = sum(level=="elementary"),
         n_sch_mid = sum(level=="middle"),
         n_sch_high = sum(level=="high"))
#now prep table
level_diffs_tbl_2 <- level_diffs %>% 
  mutate(pct_school_elementary = n_elementary / counts$n_sch_elem, #denom = total schools within category
         pct_school_middle = n_middle / counts$n_sch_mid,
         pct_school_high = n_high / counts$n_sch_high,
         elem_mid_diff = abs(pct_school_elementary - pct_school_middle), #modify percentages to reflect within groups
         mid_high_diff = abs(pct_school_middle - pct_school_high),
         high_elem_diff = abs(pct_school_high - pct_school_elementary)) %>% 
  left_join(tag_labels, by = "tag") %>% 
  select(label, elem_mid_diff, mid_high_diff, high_elem_diff, pct_school_elementary, pct_school_middle, pct_school_high, n_elementary, n_middle, n_high) %>% 
  #pretty up cols
  mutate(pct_school_elementary = paste0(round(100*pct_school_elementary, 2), "%"),
         pct_school_middle = paste0(round(100*pct_school_middle, 2), "%"),
         pct_school_high = paste0(round(100*pct_school_high, 2), "%"),
         elem_mid_diff = paste0(round(100*elem_mid_diff, 2), "%"),
         mid_high_diff = paste0(round(100*mid_high_diff, 2), "%"),
         high_elem_diff = paste0(round(100*high_elem_diff, 2), "%")) %>% 
  rename(`Tag` = label,
         `Elementary vs. Middle Difference` = elem_mid_diff,
         `Middle vs. High Difference` = mid_high_diff,
         `High vs. Elementary Difference` = high_elem_diff,
         `Pct. Tags (Elementary)` = pct_school_elementary,
         `Pct. Tags (Middle)` = pct_school_middle,
         `Pct. Tags (High)` = pct_school_high,
         `N Tags (Elementary)` = n_elementary,
         `N Tags (Middle)` = n_middle,
         `N Tags (High)` = n_high)
#table
datatable(level_diffs_tbl_2)
```

```{r within dat}
within_diffs <- level_diffs %>% 
  mutate(pct_school_elementary = n_elementary / counts$n_sch_elem, #denom = total schools within category
         pct_school_middle = n_middle / counts$n_sch_mid,
         pct_school_high = n_high / counts$n_sch_high,
         elem_mid_diff = abs(pct_school_elementary - pct_school_middle), #modify percentages to reflect within groups
         mid_high_diff = abs(pct_school_middle - pct_school_high),
         high_elem_diff = abs(pct_school_high - pct_school_elementary))
```

## Elementary vs. Middle - Largest Differences

```{r elem v mid lg, fig.dim = c(9, 9)}
#pull top 10 lg diff
elem_mid_diff <- within_diffs %>% 
  select(tag, elem_mid_diff, pct_school_elementary, pct_school_middle) %>% 
  slice_max(elem_mid_diff, n = 10, with_ties = FALSE) %>% 
    mutate(elem_mid_diff = paste0(round(100*elem_mid_diff, 2), "%"),
         pct_school_elementary = 100*pct_school_elementary,
         pct_school_middle = 100*pct_school_middle) %>% 
  pivot_longer(cols = c("pct_school_elementary", "pct_school_middle"),
               names_to = "level",
               names_prefix = "pct_school_",
               values_to = "pct")
#create labels
elem_mid_labs <- data.frame(x = c(elem_mid_diff$tag),
                            y = c(2 + elem_mid_diff$pct),
                            lab = c("", "14.63%", "", "13.73%", "", "13.16%", "12.69%", "", "12.09%", "", "", "11.61%", "", "11.22%", "", "10.69%", "", "10.63%", "", "9.73%"))
#reorder factor
elem_mid_diff$tag <- factor(elem_mid_diff$tag, levels = c(unique(paste0(elem_mid_diff$tag))))
#levels(elem_mid_diff$tag)
#side-side bar chart
elem_mid_viz <- elem_mid_diff %>% 
  ggplot(aes(tag, pct, fill = level)) +
  geom_bar(stat = "identity", position = "dodge") +
  scale_fill_manual(values = paste(transcend_cols)) +
  theme_transcend_sparse +
  geom_text(aes(x = elem_mid_labs$x, y = elem_mid_labs$y, label = elem_mid_labs$lab), color = "#EF464B", size = 5, family = "Bebas Neue") +
  theme(axis.text.x = element_text(angle = -55, hjust = 0)) +
  scale_x_tag() +
  scale_y_continuous(limits = c(0, 100)) +
  labs(x = "",
       y = "Percent of Schools Selecting Tag",
       title = "Top 10 Tags with Largest Differences between Elementary and Middle Schools",
       subtitle = "Comparison within level types - denominator = schools within level")

elem_mid_viz
```


## Middle vs. High - Largest Differences

```{r mid v high lg, fig.dim = c(9, 9)}
#pull top 10 lg diff
mid_high_diff <- within_diffs %>% 
  select(tag, mid_high_diff, pct_school_middle, pct_school_high) %>% 
  slice_max(mid_high_diff, n = 10, with_ties = FALSE) %>% 
    mutate(mid_high_diff = paste0(round(100*mid_high_diff, 2), "%"),
         pct_school_middle = 100*pct_school_middle,
         pct_school_high = 100*pct_school_high) %>% 
  pivot_longer(cols = c("pct_school_middle", "pct_school_high"),
               names_to = "level",
               names_prefix = "pct_school_",
               values_to = "pct")
#create labels
mid_high_labs <- data.frame(x = c(mid_high_diff$tag),
                            y = c(2 + mid_high_diff$pct),
                            lab = c("", "34.97%", "", "33.17%", "", "32.29%", "", "25.52%", "", "22.44%", "", "18.2%", "16.88%", "", "", "16.69%", "", "15.09%", "", "14.83%"))
#reorder factor
mid_high_diff$tag <- factor(mid_high_diff$tag, levels = c(unique(paste0(mid_high_diff$tag))))
#levels(mid_high_diff$tag)
#side-side bar chart
mid_high_viz <- mid_high_diff %>% 
  ggplot(aes(tag, pct, fill = level)) +
  geom_bar(stat = "identity", position = "dodge") +
  scale_fill_manual(values = paste(transcend_cols)) +
  theme_transcend_sparse +
  geom_text(aes(x = mid_high_labs$x, y = mid_high_labs$y, label = mid_high_labs$lab), color = "#EF464B", size = 5, family = "Bebas Neue") +
  theme(axis.text.x = element_text(angle = -55, hjust = 0)) +
  scale_x_tag() +
  scale_y_continuous(limits = c(0, 100)) +
  labs(x = "",
       y = "Percent of Schools Selecting Tag",
       title = "Top 10 Tags with Largest Differences between Middle and High Schools",
       subtitle = "Comparison within level types - denominator = schools within level")

mid_high_viz
```


## High vs. Elementary - Largest Differences

```{r high v elem lg, fig.dim = c(9, 9)}
#pull top 10 lg diff
high_elem_diff <- within_diffs %>% 
  select(tag, high_elem_diff, pct_school_high, pct_school_elementary) %>% 
  slice_max(high_elem_diff, n = 10, with_ties = FALSE) %>% 
    mutate(high_elem_diff = paste0(round(100*high_elem_diff, 2), "%"),
         pct_school_elementary = 100*pct_school_elementary,
         pct_school_high = 100*pct_school_high) %>% 
  pivot_longer(cols = c("pct_school_high", "pct_school_elementary"),
               names_to = "level",
               names_prefix = "pct_school_",
               values_to = "pct")
#create labels
high_elem_labs <- data.frame(x = c(high_elem_diff$tag),
                            y = c(2 + high_elem_diff$pct),
                            lab = c("46.02%", "", "43.27%", "", "32.52%", "", "31.94%", "", "", "29.57%", "22.99%", "", "21.87%", "", "20.82%", "", "18.65%", "", "18.49%", ""))
#reorder factor
high_elem_diff$tag <- factor(high_elem_diff$tag, levels = c(unique(paste0(high_elem_diff$tag))))
#levels(high_elem_diff$tag)
#side-side bar chart
high_elem_viz <- high_elem_diff %>% 
  ggplot(aes(tag, pct, fill = level)) +
  geom_bar(stat = "identity", position = "dodge") +
  scale_fill_manual(values = paste(transcend_cols)) +
  theme_transcend_sparse +
  geom_text(aes(x = high_elem_labs$x, y = high_elem_labs$y, label = high_elem_labs$lab), color = "#EF464B", size = 5, family = "Bebas Neue") +
  theme(axis.text.x = element_text(angle = -55, hjust = 0)) +
  scale_x_tag() +
  scale_y_continuous(limits = c(0, 100)) +
  labs(x = "",
       y = "Percent of Schools Selecting Tag",
       title = "Top 10 Tags with Largest Differences between High and Elementary Schools",
       subtitle = "Comparison within level types - denominator = schools within level")

high_elem_viz
```


## Elementary vs. Middle - Smallest Differences

```{r elem v mid sm, fig.dim = c(9, 9)}
#pull top 10 sm diff
elem_mid_diff_sm <- within_diffs %>% 
  select(tag, elem_mid_diff, pct_school_elementary, pct_school_middle) %>% 
  slice_min(elem_mid_diff, n = 10, with_ties = FALSE) %>% 
    mutate(elem_mid_diff = paste0(round(100*elem_mid_diff, 2), "%"),
         pct_school_elementary = 100*pct_school_elementary,
         pct_school_middle = 100*pct_school_middle) %>% 
  pivot_longer(cols = c("pct_school_elementary", "pct_school_middle"),
               names_to = "level",
               names_prefix = "pct_school_",
               values_to = "pct")
#create labels
elem_mid_labs_sm <- data.frame(x = c(elem_mid_diff_sm$tag),
                            y = c(2 + elem_mid_diff_sm$pct),
                            lab = c("< 1%", "", "", "< 1%", "", "< 1%", "< 1%", "", "< 1%", "", "", "< 1%", "", "< 1%", "< 1%", "", "", "< 1%", "< 1%", ""))
#reorder factor
elem_mid_diff_sm$tag <- factor(elem_mid_diff_sm$tag, levels = c(unique(paste0(elem_mid_diff_sm$tag))))
#levels(elem_mid_diff_sm$tag)
#side-side bar chart
elem_mid_viz_sm <- elem_mid_diff_sm %>% 
  ggplot(aes(tag, pct, fill = level)) +
  geom_bar(stat = "identity", position = "dodge") +
  scale_fill_manual(values = paste(transcend_cols)) +
  theme_transcend_sparse +
  geom_text(aes(x = elem_mid_labs_sm$x, y = elem_mid_labs_sm$y, label = elem_mid_labs_sm$lab), color = "#EF464B", size = 5, family = "Bebas Neue") +
  theme(axis.text.x = element_text(angle = -55, hjust = 0)) +
  scale_x_tag() +
  scale_y_continuous(limits = c(0, 100)) +
  labs(x = "",
       y = "Percent of Schools Selecting Tag",
       title = "Top 10 Tags with Smallest Differences between Elementary and Middle Schools",
       subtitle = "Comparison within level types - denominator = schools within level")

elem_mid_viz_sm
```


## Middle vs. High - Smallest Differences

```{r mid v high sm, fig.dim = c(9, 9)}
#pull top 10 lg diff
mid_high_diff_sm <- within_diffs %>% 
  select(tag, mid_high_diff, pct_school_middle, pct_school_high) %>% 
  slice_min(mid_high_diff, n = 10, with_ties = FALSE) %>% 
    mutate(mid_high_diff = paste0(round(100*mid_high_diff, 2), "%"),
         pct_school_middle = 100*pct_school_middle,
         pct_school_high = 100*pct_school_high) %>% 
  pivot_longer(cols = c("pct_school_middle", "pct_school_high"),
               names_to = "level",
               names_prefix = "pct_school_",
               values_to = "pct")
#create labels
mid_high_labs_sm <- data.frame(x = c(mid_high_diff_sm$tag),
                            y = c(4 + mid_high_diff_sm$pct),
                            lab = c("", "< 1%", "<1 %", "", "< 1%", "", "< 1%", "", "", "< 1%", "< 1%", "", "", "< 1%", "", "1%", "", "1.06%", "", "1.35%"))
#reorder factor
mid_high_diff_sm$tag <- factor(mid_high_diff_sm$tag, levels = c(unique(paste0(mid_high_diff_sm$tag))))
#levels(mid_high_diff_sm$tag)
#side-side bar chart
mid_high_viz_sm <- mid_high_diff_sm %>% 
  ggplot(aes(tag, pct, fill = level)) +
  geom_bar(stat = "identity", position = "dodge") +
  scale_fill_manual(values = paste(transcend_cols)) +
  theme_transcend_sparse +
  geom_text(aes(x = mid_high_labs_sm$x, y = mid_high_labs_sm$y, label = mid_high_labs_sm$lab), color = "#EF464B", size = 5, family = "Bebas Neue") +
  theme(axis.text.x = element_text(angle = -55, hjust = 0)) +
  scale_x_tag() +
  scale_y_continuous(limits = c(0, 100)) +
  labs(x = "",
       y = "Percent of Schools Selecting Tag",
       title = "Top 10 Tags with Smallest Differences between Middle and High Schools",
       subtitle = "Comparison within level types - denominator = schools within level")

mid_high_viz_sm
```


## High vs. Elementary - Smallest Differences

```{r high v elem sm, fig.dim = c(9, 9)}
#pull top 10 lg diff
high_elem_diff_sm <- within_diffs %>% 
  select(tag, high_elem_diff, pct_school_high, pct_school_elementary) %>% 
  slice_min(high_elem_diff, n = 10, with_ties = FALSE) %>% 
    mutate(high_elem_diff = paste0(round(100*high_elem_diff, 2), "%"),
         pct_school_elementary = 100*pct_school_elementary,
         pct_school_high = 100*pct_school_high) %>% 
  pivot_longer(cols = c("pct_school_high", "pct_school_elementary"),
               names_to = "level",
               names_prefix = "pct_school_",
               values_to = "pct")
#create labels
high_elem_labs_sm <- data.frame(x = c(high_elem_diff_sm$tag),
                            y = c(2 + high_elem_diff_sm$pct),
                            lab = c("", "< 1%", "< 1%", "", "< 1%", "", "", "< 1%", "", "< 1%", "< 1%", "", "< 1%", "", "", "< 1%", "", "< 1%", "1.14%", ""))
#reorder factor
high_elem_diff_sm$tag <- factor(high_elem_diff_sm$tag, levels = c(unique(paste0(high_elem_diff_sm$tag))))
#levels(high_elem_diff_sm$tag)
#side-side bar chart
high_elem_viz_sm <- high_elem_diff_sm %>% 
  ggplot(aes(tag, pct, fill = level)) +
  geom_bar(stat = "identity", position = "dodge") +
  scale_fill_manual(values = paste(transcend_cols)) +
  theme_transcend_sparse +
  geom_text(aes(x = high_elem_labs_sm$x, y = high_elem_labs_sm$y, label = high_elem_labs_sm$lab), color = "#EF464B", size = 5, family = "Bebas Neue") +
  theme(axis.text.x = element_text(angle = -55, hjust = 0)) +
  scale_x_tag() +
  scale_y_continuous(limits = c(0, 100)) +
  labs(x = "",
       y = "Percent of Schools Selecting Tag",
       title = "Top 10 Tags with Smallest Differences between High and Elementary Schools",
       subtitle = "Comparison within level types - denominator = schools within level")

high_elem_viz_sm
```

Save plots:
```{r, eval = FALSE}
#largest diffs
ggsave("elem-mid_lg-diff.png", plot = elem_mid_viz, path = here("outputs", "level analysis"),
       width = 12, height = 8, units = "in")
ggsave("mid-high_lg-diff.png", plot = mid_high_viz, path = here("outputs", "level analysis"),
       width = 12, height = 8, units = "in")
ggsave("high-elem_lg-diff.png", plot = high_elem_viz, path = here("outputs", "level analysis"),
       width = 12, height = 8, units = "in")
#smallest diffs
ggsave("elem-mid_sm-diff.png", plot = elem_mid_viz_sm, path = here("outputs", "level analysis"),
       width = 12, height = 8, units = "in")
ggsave("mid-high_sm-diff.png", plot = mid_high_viz_sm, path = here("outputs", "level analysis"),
       width = 12, height = 8, units = "in")
ggsave("high-elem_sm-diff.png", plot = high_elem_viz_sm, path = here("outputs", "level analysis"),
       width = 12, height = 8, units = "in")
```

# Variation Across All Levels

```{r all variability, fig.dim = c(9, 9)}
#data prep
#variance calculated as standard deviation within tag groups
var_all <- level_diffs %>% 
  select(-ends_with("_diff")) %>% 
  mutate(pct_school_elementary = n_elementary / counts$n_sch_elem, #denom = total schools within category
         pct_school_middle = n_middle / counts$n_sch_mid,
         pct_school_high = n_high / counts$n_sch_high) %>% 
  pivot_longer(cols = c("pct_school_elementary", "pct_school_middle", "pct_school_high"),
               names_to = "level",
               names_prefix = "pct_school_",
               values_to = "proportion") %>% 
  ungroup() %>% 
  group_by(tag) %>% 
  mutate(variance = sd(proportion))

#max variability plot - top 10
var_max_10 <- var_all %>% 
  group_by(level) %>% 
  slice_max(variance, n = 10, with_ties = FALSE) %>% 
  ggplot(aes(tag, proportion, fill = level)) + 
    geom_bar(stat = "identity", position = "dodge") +
  scale_fill_manual(values = paste(transcend_cols)) +
  theme_transcend_sparse +
  #geom_text(aes(x = tag, y = ind_labs$y, label = ind_labs$lab),
  #          color = "#EF464B", size = 5, family = "Bebas Neue") +
  theme(axis.text.x = element_text(angle = -55, hjust = 0)) +
  #scale_y_continuous(limits = c(0, 100)) +
  scale_x_tag() +
  labs(x = "",
       y = "Tags with Most Variability",
       title = "Top 10 Tags with Largest Variability for Elementary, Middle, and High Schools",
       subtitle = "In order of increasing variability")

var_max_10

#min variability plot - top 10
var_min_10 <- var_all %>% 
  group_by(level) %>% 
  slice_min(variance, n = 10, with_ties = FALSE) %>% 
  ggplot(aes(tag, proportion, fill = level)) + 
    geom_bar(stat = "identity", position = "dodge") +
  scale_fill_manual(values = paste(transcend_cols)) +
  theme_transcend_sparse +
  #geom_text(aes(x = tag, y = ind_labs$y, label = ind_labs$lab),
  #          color = "#EF464B", size = 5, family = "Bebas Neue") +
  theme(axis.text.x = element_text(angle = -55, hjust = 0)) +
  #scale_y_continuous(limits = c(0, 100)) +
  scale_x_tag() +
  labs(x = "",
       y = "Tags with Least Variability",
       title = "Lowest 10 Tags with Largest Variability for Elementary, Middle, and High Schools",
       subtitle = "In order of increasing variability")

var_min_10

#max variability plot - top 15
var_max_15 <- var_all %>% 
  group_by(level) %>% 
  slice_max(variance, n = 15, with_ties = FALSE) %>% 
  ggplot(aes(tag, proportion, fill = level)) + 
    geom_bar(stat = "identity", position = "dodge") +
  scale_fill_manual(values = paste(transcend_cols)) +
  theme_transcend_sparse +
  #geom_text(aes(x = tag, y = ind_labs$y, label = ind_labs$lab),
  #          color = "#EF464B", size = 5, family = "Bebas Neue") +
  theme(axis.text.x = element_text(angle = -55, hjust = 0)) +
  #scale_y_continuous(limits = c(0, 100)) +
  scale_x_tag() +
  labs(x = "",
       y = "Tags with Most Variability",
       title = "Top 15 Tags with Largest Variability for Elementary, Middle, and High Schools",
       subtitle = "In order of increasing variability")

var_max_15

#min variability plot - top 15
var_min_15 <- var_all %>% 
  group_by(level) %>% 
  slice_min(variance, n = 15, with_ties = FALSE) %>% 
  ggplot(aes(tag, proportion, fill = level)) + 
    geom_bar(stat = "identity", position = "dodge") +
  scale_fill_manual(values = paste(transcend_cols)) +
  theme_transcend_sparse +
  #geom_text(aes(x = tag, y = ind_labs$y, label = ind_labs$lab),
  #          color = "#EF464B", size = 5, family = "Bebas Neue") +
  theme(axis.text.x = element_text(angle = -55, hjust = 0)) +
  #scale_y_continuous(limits = c(0, 100)) +
  scale_x_tag() +
  labs(x = "",
       y = "Tags with Least Variability",
       title = "Lowest 15 Tags with Largest Variability for Elementary, Middle, and High Schools",
       subtitle = "In order of increasing variability")

var_min_15

#max variability plot - top 20
var_max_20 <- var_all %>% 
  group_by(level) %>% 
  slice_max(variance, n = 20, with_ties = FALSE) %>% 
  ggplot(aes(tag, proportion, fill = level)) + 
    geom_bar(stat = "identity", position = "dodge") +
  scale_fill_manual(values = paste(transcend_cols)) +
  theme_transcend_sparse +
  #geom_text(aes(x = tag, y = ind_labs$y, label = ind_labs$lab),
  #          color = "#EF464B", size = 5, family = "Bebas Neue") +
  theme(axis.text.x = element_text(angle = -55, hjust = 0)) +
  #scale_y_continuous(limits = c(0, 100)) +
  scale_x_tag() +
  labs(x = "",
       y = "Tags with Most Variability",
       title = "Top 20 Tags with Largest Variability for Elementary, Middle, and High Schools",
       subtitle = "In order of increasing variability")

var_max_20

#min variability plot - top 20
var_min_20 <- var_all %>% 
  group_by(level) %>% 
  slice_min(variance, n = 20, with_ties = FALSE) %>% 
  ggplot(aes(tag, proportion, fill = level)) + 
    geom_bar(stat = "identity", position = "dodge") +
  scale_fill_manual(values = paste(transcend_cols)) +
  theme_transcend_sparse +
  #geom_text(aes(x = tag, y = ind_labs$y, label = ind_labs$lab),
  #          color = "#EF464B", size = 5, family = "Bebas Neue") +
  theme(axis.text.x = element_text(angle = -55, hjust = 0)) +
  #scale_y_continuous(limits = c(0, 100)) +
  scale_x_tag() +
  labs(x = "",
       y = "Tags with Least Variability",
       title = "Lowest 20 Tags with Largest Variability for Elementary, Middle, and High Schools",
       subtitle = "In order of increasing variability")

var_min_20
```

Save plots

```{r save variance plots, eval = FALSE}
#variability across tags
#top 10
ggsave("lg-tag-variability-10.png", plot = var_max_10, path = here("outputs", "level analysis"),
       width = 12, height = 8, units = "in")
ggsave("sm-tag-variability-10.png", plot = var_min_10, path = here("outputs", "level analysis"),
       width = 12, height = 8, units = "in")
#top 15
ggsave("lg-tag-variability-15.png", plot = var_max_15, path = here("outputs", "level analysis"),
       width = 12, height = 8, units = "in")
ggsave("sm-tag-variability-15.png", plot = var_min_15, path = here("outputs", "level analysis"),
       width = 12, height = 8, units = "in")
#top 20
ggsave("lg-tag-variability-20.png", plot = var_max_20, path = here("outputs", "level analysis"),
       width = 12, height = 8, units = "in")
ggsave("sm-tag-variability-20.png", plot = var_min_20, path = here("outputs", "level analysis"),
       width = 12, height = 8, units = "in")
```

