---
title: "Level Analysis"
author: "Janette Avelar"
date: "5/23/2022"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
```

```{r library}
library(here)
library(rio)
library(psych)
library(tidyverse)
library(ggplot2)
library(DT)
library(scales)
source("scripts/branding.R", local = knitr::knit_global())
load(here("data", "complete_canopy_2022.RData"))
```

# Cycle 4 (5.24.22) [in preparation for Brown Bag]

Guiding questions/tasks & format:  
`LEVEL`: How do tagging patterns differ among elementary, middle, and high schools?  
* table of biggest differences between categories (elem vs. middle; middle vs. high; elem vs. high)  
* odds ratio  
* visualization  

## Analysis Notes:  
* elementary schools = 50  
* middle schools = 67  
* high schools = 93  
* elementary AND middle = 38  
* middle AND high = 32  
* elementary AND middle AND high = 19  
* NONE elementary middle or high = 11  
    * CodeRVA Regional High School   
    * Empower Community High School  
    * Future Public School  
    * Purdue Polytechnic High School  
    * Rooted School  
    * Living School  
    * Achieve Community Charter School  
    * Southeast Area Technical High School (Sea-Tech)  
    * Sequoyah School  
    * Vimenti School  
    * Foundation School Of Autism - San Antonio  
* 10 `NA` values
* total schools = 161 - 10 `NA` = 151


*Focus on odds ratio plot for levels by Tuesday--other plots by this weekend*

```{r data set up and clean, include = FALSE}
level_tags <- clean_data_full %>% 
  select(school_id, grades_elementary, grades_middle, grades_high, starts_with("practices")) %>% 
  filter(!c(is.na(grades_elementary) & is.na(grades_middle) & is.na(grades_high))) %>% 
  mutate(grades_elementary = ifelse(grades_elementary==1, "elementary", "none"),
         grades_middle = ifelse(grades_middle==1, "middle", "none"),
         grades_high = ifelse(grades_high==1, "high", "none")) %>% 
  pivot_longer(cols = starts_with("grades"),
               names_to = "og_level",
               values_to = "level") %>% 
  select(-og_level) %>% 
  filter(!level=="none") %>% 
  pivot_longer(cols = starts_with("practices"),
               names_to = "tag",
               values_to = "n") %>% 
  select(-school_id) %>% 
  group_by(level, tag) %>% 
  mutate(n = sum(n, na.rm = TRUE)) %>% 
  ungroup() %>% 
  unique() %>% 
#percents for each practice by level
  group_by(level) %>% 
  mutate(total_tags_level = sum(n),
         pct_school = n/151, #only 151 schools in the analysis
         pct_total_tags = n/total_tags_level) #schools in each locale per tag / total tags in locale - ignore for odds
```

## Probability and Odds Ratios

In this final analysis, I calculated probability for each practice and converted to odds ratio in order to plot how much more or less likely a given locale was to select each tag.

```{r level odds}
level_odds <- level_tags %>% 
  select(level, tag, pct_school) %>% 
  pivot_wider(names_from = "level",
              values_from = "pct_school",
              names_glue = "{.value}_{level}") %>% 
  janitor::clean_names() %>% 
  mutate(odds_elem = pct_school_elementary / (1 - pct_school_elementary),
         odds_mid = pct_school_middle / (1 - pct_school_middle),
         odds_high = pct_school_high / (1 - pct_school_high),
         elem_mid_ratio = odds_elem / odds_mid,
         mid_high_ratio = odds_mid / odds_high,
         high_elem_ratio = odds_high / odds_elem)
```

### Elementary vs. Middle (odds ratio)

```{r elem v mid odd viz, fig.dim = c(9, 9)}
elem_mid_odds <- level_odds %>% 
  select(tag, elem_mid_ratio) %>% 
  slice_max(elem_mid_ratio, n = 15, with_ties = FALSE)
#odd ratio labels
elem_mid_odds_lab <- data.frame(x = c(elem_mid_odds$tag),
                           y = c(elem_mid_odds$elem_mid_ratio + .05),
                           lab = c(paste0(round(elem_mid_odds$elem_mid_ratio, 1), "x")))
#viz
elem_mid_odds_viz <- elem_mid_odds %>% 
  ggplot(aes(reorder(tag, -elem_mid_ratio), elem_mid_ratio)) +
  geom_bar(stat = "identity", position = "dodge", fill = "#1A4C81") +
  coord_flip() +
  geom_text(aes(x = elem_mid_odds_lab$x, y = elem_mid_odds_lab$y, label = elem_mid_odds_lab$lab), 
            color = "#EF464B", size = 7, family = "Bebas Neue") +
  scale_color_manual(values = paste(transcend_cols)) +
  scale_x_tag() +
  theme_transcend +
  theme(axis.text.x = element_blank(),
        axis.ticks.x = element_blank()) +
  labs(x = "",
       y = "",
       title = "How Much More Likely Elementary Schools were to \nSelect a Tag",
       subtitle = "compared to middle schools using raw odds ratio")
  
elem_mid_odds_viz
```

### Middle vs. High (odds ratio)

```{r elem v mid odd viz, fig.dim = c(9, 9)}
mid_high_odds <- level_odds %>% 
  select(tag, mid_high_ratio) %>% 
  slice_max(mid_high_ratio, n = 15, with_ties = FALSE)
#odd ratio labels
mid_high_odds_lab <- data.frame(x = c(mid_high_odds$tag),
                           y = c(mid_high_odds$mid_high_ratio + .05),
                           lab = c(paste0(round(mid_high_odds$mid_high_ratio, 1), "x")))
#viz
mid_high_odds_viz <- mid_high_odds %>% 
  ggplot(aes(reorder(tag, -mid_high_ratio), mid_high_ratio)) +
  geom_bar(stat = "identity", position = "dodge", fill = "#1A4C81") +
  coord_flip() +
  geom_text(aes(x = mid_high_odds_lab$x, y = mid_high_odds_lab$y, label = mid_high_odds_lab$lab), 
            color = "#EF464B", size = 7, family = "Bebas Neue") +
  scale_color_manual(values = paste(transcend_cols)) +
  scale_x_tag() +
  theme_transcend +
  theme(axis.text.x = element_blank(),
        axis.ticks.x = element_blank()) +
  labs(x = "",
       y = "",
       title = "How Much More Likely Middle Schools were to \nSelect a Tag",
       subtitle = "compared to high schools using raw odds ratio")
  
mid_high_odds_viz
```

### High vs. Elementary (odds ratio)

```{r elem v mid odd viz, fig.dim = c(9, 9)}
high_elem_odds <- level_odds %>% 
  select(tag, high_elem_ratio) %>% 
  slice_max(high_elem_ratio, n = 15, with_ties = FALSE)
#odd ratio labels
high_elem_odds_lab <- data.frame(x = c(high_elem_odds$tag),
                           y = c(high_elem_odds$high_elem_ratio + .35),
                           lab = c(paste0(round(high_elem_odds$high_elem_ratio, 1), "x")))
#viz
high_elem_odds_viz <- high_elem_odds %>% 
  ggplot(aes(reorder(tag, -high_elem_ratio), high_elem_ratio)) +
  geom_bar(stat = "identity", position = "dodge", fill = "#1A4C81") +
  coord_flip() +
  geom_text(aes(x = high_elem_odds_lab$x, y = high_elem_odds_lab$y, label = high_elem_odds_lab$lab), 
            color = "#EF464B", size = 7, family = "Bebas Neue") +
  scale_color_manual(values = paste(transcend_cols)) +
  scale_x_tag() +
  theme_transcend +
  theme(axis.text.x = element_blank(),
        axis.ticks.x = element_blank()) +
  labs(x = "",
       y = "",
       title = "How Much More Likely High Schools were to \nSelect a Tag",
       subtitle = "compared to elementary schools using raw odds ratio")
  
high_elem_odds_viz
```

Save plots:
```{r, eval = FALSE}
#odds charts
ggsave("elem-mid_odds_chart_raw.png", plot = elem_mid_odds_viz, path = here("outputs", "level analysis"),
       width = 12, height = 8, units = "in")
ggsave("mid-high_odds_chart_raw.png", plot = mid_high_odds_viz, path = here("outputs", "level analysis"),
       width = 12, height = 8, units = "in")
ggsave("high-elem_odds_chart_raw.png", plot = high_elem_odds_viz, path = here("outputs", "level analysis"),
       width = 12, height = 8, units = "in")
```