---
title: "Locale Analysis"
author: "Janette Avelar"
date: "5/18/2022"
output: 
  html_document:
    toc: true
    toc_float: true
    theme: flatly
    highlight: tango
    code_folding: hide
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
```

```{r library, include = FALSE}
library(here)
library(rio)
library(psych)
library(tidyverse)
library(ggplot2)
library(DT)
library(scales)
source("scripts/branding.R", local = knitr::knit_global())
```

```{r modify practices data}
load(here("data", "complete_canopy_2022.RData"))
tags <- import(here("data", "Tags-SY2021-22 Tags.csv"))
```


# Cycle 3 (5.18.22)

Guiding questions/tasks & format:  
`LOCALE`: How do tagging patterns differ among rural, urban, and suburban schools?  
* table of biggest differences between categories (urban vs. rural; rural vs. suburban; suburban vs. urban)  
* odds ratio  
* visualization  

## Analysis Notes 

Data used:  
There are 4 schools not included in analysis:  
* Horizons Alternative (no NCES/self-reported data)  
* Iowa Big (multi-coded locale)  
* The Forest Schools (multi-coded locale)  
* Link Learning (multi-coded locale)  

Variables created:  
* `n` - number of schools that selected tag  
* `total_tags_locale` - total number of tags in given locale category  
* `pct_school` - percentage of schools that selected tag (n/157)  
* `pct_total_tags` - percentage of total tags within each locale (n/total_tags_locale)

```{r data set up and clean, include = FALSE}
#create totals for each practice by locale
locale_tags <- clean_data_full %>% 
  select(school_id, locale) %>% 
  left_join(practices_data, by = "school_id") %>% 
  select(school_id, locale, starts_with("practices_")) %>% 
  filter(!c(locale == "")) %>%  #Horizons Alternative; Iowa Big; The Forest Schools; Link Learning
  # mutate_at(vars(starts_with("practices_")), as.integer, na.rm = TRUE) %>% 
  pivot_longer(cols = starts_with("practices"),
               names_to = "tag",
               values_to = "n") %>% 
  select(-c(school_id)) %>% 
  group_by(locale, tag) %>% 
  mutate(n = sum(n, na.rm = TRUE)) %>% 
  unique() %>% 
#percents for each practice by locale
  ungroup() %>% 
  group_by(locale) %>% 
  mutate(total_tags_locale = sum(n),
         pct_school = n/157, #only 157 schools in the analysis
         pct_total_tags = n/total_tags_locale) #schools in each locale per tag / total tags in locale
```

## Across Locale Table and Plots

The following table and plots present the data within locales as a function of the full dataset.  

In other words, the denominator is 157 (161 total schools - 4 schools excluded from analysis either because of either missing or multi-coded data).  

As a result, the bar charts are perhaps not easily compared and/or what we were aiming for, something I realized after creating them, but I left them in for us to look at in case there's any other information they help give us, such as overall distribution of tags.  

*Note:* I haven't incorporated theming into the plots yet--I need to figure out how to download fonts and work with theming, which I am not super familiar with and ran out of time.

### Table - Tagging Differences by Locale (across all schools)

```{r diff tbl}
#create difference columns
locale_diffs <- locale_tags %>% 
  select(locale, tag, n, pct_school) %>% 
  pivot_wider(names_from = "locale",
              values_from = c("n", "pct_school")) %>% 
  janitor::clean_names() %>% 
  mutate(urb_rur_diff = abs(round(100*(pct_school_urban - pct_school_rural), 2)), #diff b/w urban & rural
         urb_sub_diff = abs(round(100*(pct_school_urban - pct_school_suburban), 2)), #diff b/w urban & suburban
         sub_rur_diff = abs(round(100*(pct_school_suburban - pct_school_rural), 2))) #diff b/w suburban & rural
#prepare data for table with nice labels
locale_diffs_tbl <- tags %>% 
  janitor::clean_names() %>% 
  select(variable_name, short_name) %>%
  rename(tag = variable_name) %>% 
  left_join(locale_diffs, by = "tag") %>% 
  select(short_name, urb_rur_diff, urb_sub_diff, sub_rur_diff,
         pct_school_urban, pct_school_suburban, pct_school_rural,
         n_urban, n_suburban, n_rural) %>% 
  mutate(urb_rur_diff = paste(round(1*urb_rur_diff, 0), "%"),
         urb_sub_diff = paste(round(1*urb_sub_diff, 0), "%"),
         sub_rur_diff = paste(round(1*sub_rur_diff, 0), "%"),
         pct_school_urban = paste(round(100*pct_school_urban, 0), "%"),
         pct_school_suburban = paste(round(100*pct_school_suburban, 0), "%"),
         pct_school_rural = paste(round(100*pct_school_rural, 0), "%")) %>% 
  rename(`Tag` = short_name,
         `Urban vs. Rural Difference` =  urb_rur_diff,
         `Urban vs. Suburban Difference` = urb_sub_diff,
         `Suburban vs. Rural Difference` = sub_rur_diff,
         `Pct. Tags (Urban)` = pct_school_urban,
         `Pct. Tags (Suburban)` = pct_school_suburban,
         `Pct. Tags (Rural)` = pct_school_rural,
         `N Tags (Urban)` = n_urban,
         `N Tags (Suburban)` = n_suburban,
         `N Tags (Rural)` = n_rural)
#table
datatable(locale_diffs_tbl)
```

### Urban vs. Rural Barchart (Across)  

```{r urb v rur viz, fig.dim = c(9, 9)}
#source("scripts/branding.R")
#prep data for viz
urb_rur_diff <- tags %>% 
  janitor::clean_names() %>% 
  select(variable_name, short_name) %>% 
  rename(tag = variable_name) %>% 
  left_join(locale_diffs, by = "tag") %>% 
  select(short_name, urb_rur_diff, pct_school_urban, pct_school_rural) %>% 
  ungroup() %>% 
  slice_max(urb_rur_diff, n = 10, with_ties = FALSE) %>% 
  mutate(urb_rur_diff = paste(round(1*urb_rur_diff, 0), "%"),
         pct_school_urban = 100*pct_school_urban,
         pct_school_rural = 100*pct_school_rural) %>% 
  pivot_longer(cols = c("pct_school_urban", "pct_school_rural"),
               names_to = "locale",
               names_prefix = "pct_school_",
               values_to = "pct")
#pct labels
urb_rur_labs <- data.frame(x = c(urb_rur_diff$short_name),
                           y = c(5 + urb_rur_diff$pct),
                           lab = c("40%", "", "36%", "", "35%", "", "34%", "", "34%", "", "33%", "", "32%", "", "32%", "", "31%", "", "31%", ""))
#reorder tags
urb_rur_diff$short_name <- factor(urb_rur_diff$short_name,
                                    levels = c(unique(paste0(urb_rur_diff$short_name))))
#side-side bar chart
urb_rur_viz <- urb_rur_diff %>% 
  ggplot(aes(short_name, pct, fill = locale)) +
  geom_bar(stat = "identity", position = "dodge") +
  scale_fill_manual(values = paste(transcend_cols)) +
  geom_text(aes(x = short_name, y = urb_rur_labs$y, label = urb_rur_labs$lab)) +
  theme_transcend_sparse +
  theme(axis.text.x = element_text(angle = -55, hjust = 0)) +
  scale_y_continuous(limits = c(0, 70)) +
  labs(x = "",
       y = "Percent of Schools Selecting Tag",
       title = "Top 10 Tags with Largest Differences between Urban and Rural Schools",
       subtitle = "Comparison across locale types - denominator = total schools")

urb_rur_viz
```

### Urban vs. Suburban Barchart (Across)

```{r urb v sub viz, fig.dim = c(9, 9)}
urb_sub_diff <- tags %>% 
  janitor::clean_names() %>% 
  select(variable_name, short_name) %>% 
  rename(tag = variable_name) %>% 
  left_join(locale_diffs, by = "tag") %>% 
  select(short_name, urb_sub_diff, pct_school_urban, pct_school_suburban) %>% 
  slice_max(urb_sub_diff, n = 10, with_ties = FALSE) %>% 
  mutate(urb_sub_diff = paste(round(1*urb_sub_diff, 0), "%"),
         pct_school_urban = 100*pct_school_urban,
         pct_school_suburban = 100*pct_school_suburban) %>% 
  pivot_longer(cols = c("pct_school_urban", "pct_school_suburban"),
               names_to = "locale",
               names_prefix = "pct_school_",
               values_to = "pct")
#pct labels
urb_sub_labs <- data.frame(x = c(urb_sub_diff$short_name),
                           y = c(5 + urb_sub_diff$pct),
                           lab = c("34%", "", "32%", "", "32%", "", "32%", "", "31%", "", "31%", "", "31%", "", "31%", "", "29%", "", "29%", ""))
#reorder tags
urb_sub_diff$short_name <- factor(urb_sub_diff$short_name,
                                    levels = c(unique(paste0(urb_sub_diff$short_name))))
#side-side bar chart
urb_sub_viz <- urb_sub_diff %>% 
  ggplot(aes(short_name, pct, fill = locale)) +
  geom_bar(stat = "identity", position = "dodge") +
  ylim(0, 70) +
  scale_fill_manual(values = paste(transcend_cols)) +
  geom_text(aes(x = short_name, y = urb_sub_labs$y, label = urb_sub_labs$lab)) +
  theme_transcend_sparse +
  theme(axis.text.x = element_text(angle = -55, hjust = 0)) +
  labs(x = "",
       y = "Percent of Schools Selecting Tag",
       title = "Top 10 Tags with Largest Differences between Urban and Suburban Schools",
       subtitle = "Comparison across locale types - denominator = total schools")

urb_sub_viz
```

### Suburban vs. Rural Barchart (Across)

```{r sub v rur viz, fig.dim = c(9, 9)}
sub_rur_diff <- tags %>% 
  janitor::clean_names() %>% 
  select(variable_name, short_name) %>% 
  rename(tag = variable_name) %>% 
  left_join(locale_diffs, by = "tag") %>% 
  select(short_name, sub_rur_diff, pct_school_rural, pct_school_suburban) %>% 
  slice_max(sub_rur_diff, n = 10, with_ties = FALSE) %>% 
  mutate(sub_rur_diff = paste(round(1*sub_rur_diff, 0), "%"),
         pct_school_rural = 100*pct_school_rural,
         pct_school_suburban = 100*pct_school_suburban) %>% 
  pivot_longer(cols = c("pct_school_rural", "pct_school_suburban"),
               names_to = "locale",
               names_prefix = "pct_school_",
               values_to = "pct")
#pct labels
sub_rur_labs <- data.frame(x = c(sub_rur_diff$short_name),
                           y = c(1 + sub_rur_diff$pct),
                           lab = c("", "8%", "", "6%", "", "6%", "", "6%", "", "6%", "", "6%", "", "6%", "", "5%", "", "5%", "", "5%"))
#reorder tags
sub_rur_diff$short_name <- factor(sub_rur_diff$short_name,
                                    levels = c(unique(paste0(sub_rur_diff$short_name))))
#side-side bar chart
sub_rur_viz <- sub_rur_diff %>% 
  ggplot(aes(short_name, pct, fill = locale)) +
  geom_bar(stat = "identity", position = "dodge") +
  ylim(0, 30) +
  scale_fill_manual(values = paste(transcend_cols)) +
  geom_text(aes(x = short_name, y = sub_rur_labs$y, label = sub_rur_labs$lab)) +
  theme_transcend_sparse +
  theme(axis.text.x = element_text(angle = -55, hjust = 0)) +
  labs(x = "",
       y = "Percent of Schools Selecting Tag",
       title = "Top 10 Tags with Largest Differences between Rural and Suburban Schools",
       subtitle = "Comparison across locale types - denominator = total schools")

sub_rur_viz
```

## Within Locale Table and Plots

In my second shot at plots, I've used tagging within locale as the basis for comparison. In other words, the denominator is equal to the number of schools within the specified locale type (95 urban schools, 32 suburban schools, and 30 rural schools = 157 total schools in analysis). This gives us an idea of the differences in tagging patterns within urban, suburban, and rural areas respectively, but I was hesitant to compare due to differences in sample size which is why I plan to also plot odds ratios in the third iteration of plotting below. Still useful to consider questions like, how much more often did suburban areas implement `x` practice than rural areas? 

```{r data set up and clean 2, include = FALSE}
#create totals for each practice by locale
locale_tags_2 <- clean_data_full %>% 
  select(school_id, school_name, locale) %>% 
  left_join(practices_data, by = "school_id") %>% 
  select(school_id, school_name, locale, starts_with("practices_")) %>% 
  filter(!c(locale == "")) %>%  #Horizons Alternative; Iowa Big; The Forest Schools; Link Learning
  group_by(locale) %>% 
  mutate(n_school = length(school_name)) %>% 
  mutate_at(vars(starts_with("practices_")), as.integer, na.rm = TRUE) %>% 
  pivot_longer(cols = starts_with("practices"),
               names_to = "tag",
               values_to = "n_tag") %>% 
  select(-c(school_id, school_name)) %>% 
  ungroup() %>% 
  group_by(locale, tag) %>% 
  mutate(n_tag = sum(n_tag, na.rm = TRUE)) %>% 
  unique() %>% 
#percents for each practice by locale
  ungroup() %>% 
  group_by(locale) %>% 
  mutate(pct_locale_tag = n_tag / n_school) #total tags for given practice within locale / total school n within locale
```

### Within Locale Table

Table mainly to show the differences the within locale analysis makes to help conceptualize. (i.e., I needed this table to think through it. Notice Pct. Tags sum in this table = 100% meaning we're accounting for all tags within each practice and analyzing the distribution across locale within it, as compared to earlier table where Pct. Tags *does not* equal 100, and reflects percentage of total schools that selected said tag as distributed across locale.)

```{r within diff tbl}
#create difference columns
locale_diffs_2 <- locale_tags_2 %>% 
  select(locale, tag, n_tag, pct_locale_tag) %>% 
  pivot_wider(names_from = "locale",
              values_from = c("n_tag", "pct_locale_tag")) %>% 
  janitor::clean_names() %>% 
  mutate(urb_rur_diff = abs(round(100*(pct_locale_tag_urban - pct_locale_tag_rural), 2)), #diff b/w urban & rural
         urb_sub_diff = abs(round(100*(pct_locale_tag_urban - pct_locale_tag_suburban), 2)), #diff b/w urban & suburban
         sub_rur_diff = abs(round(100*(pct_locale_tag_suburban - pct_locale_tag_rural), 2))) #diff b/w suburban & rural
#prepare data for table with nice labels
locale_diffs_tbl_2 <- tags %>% 
  janitor::clean_names() %>% 
  select(variable_name, short_name) %>%
  rename(tag = variable_name) %>% 
  left_join(locale_diffs_2, by = "tag") %>% 
  select(short_name, urb_rur_diff, urb_sub_diff, sub_rur_diff,
         pct_locale_tag_urban, pct_locale_tag_suburban, pct_locale_tag_rural,
         n_tag_urban, n_tag_suburban, n_tag_rural) %>% 
  mutate(urb_rur_diff = paste(round(1*urb_rur_diff, 0), "%"),
         urb_sub_diff = paste(round(1*urb_sub_diff, 0), "%"),
         sub_rur_diff = paste(round(1*sub_rur_diff, 0), "%"),
         pct_locale_tag_urban = paste(round(100*pct_locale_tag_urban, 0), "%"),
         pct_locale_tag_suburban = paste(round(100*pct_locale_tag_suburban, 0), "%"),
         pct_locale_tag_rural = paste(round(100*pct_locale_tag_rural, 0), "%")) %>% 
  rename(`Tag` = short_name,
         `Urban vs. Rural Difference` =  urb_rur_diff,
         `Urban vs. Suburban Difference` = urb_sub_diff,
         `Suburban vs. Rural Difference` = sub_rur_diff,
         `Pct. Tags (Urban)` = pct_locale_tag_urban,
         `Pct. Tags (Suburban)` = pct_locale_tag_suburban,
         `Pct. Tags (Rural)` = pct_locale_tag_rural,
         `N Tags (Urban)` = n_tag_urban,
         `N Tags (Suburban)` = n_tag_suburban,
         `N Tags (Rural)` = n_tag_rural)
#table
datatable(locale_diffs_tbl_2)
```

### Urban vs. Rural (Within)  

The biggest difference was shocking to me--68% of urban schools tagged anti-racist practices, compared to only 20% of rural schools (!!). Really curious to see how these top 3 (anti-racist practices, culturally-responsive practices, and hiring for equity and inclusion values) have or have not shifted from previous years.

```{r urb v rur viz 2, fig.dim = c(9, 9)}
#prep data
urb_rur_diff_2 <- tags %>% 
  janitor::clean_names() %>% 
  select(variable_name, short_name) %>% 
  rename(tag = variable_name) %>% 
  left_join(locale_diffs_2, by = "tag") %>% 
  select(short_name, urb_rur_diff, pct_locale_tag_urban, pct_locale_tag_rural) %>% 
  ungroup() %>% 
  slice_max(urb_rur_diff, n = 10, with_ties = FALSE) %>% 
  mutate(urb_rur_diff = paste(round(1*urb_rur_diff, 0), "%"),
         pct_locale_tag_urban = 100*pct_locale_tag_urban,
         pct_locale_tag_rural = 100*pct_locale_tag_rural) %>% 
  pivot_longer(cols = c("pct_locale_tag_urban", "pct_locale_tag_rural"),
               names_to = "locale",
               names_prefix = "pct_locale_tag_",
               values_to = "pct")
#pct labels
urb_rur_labs_2 <- data.frame(x = c(urb_rur_diff_2$short_name),
                             y = c(5 + urb_rur_diff_2$pct),
                             lab = c("49%", "", "47%", "", "42%", "", "38%", "", "29%", "", "", "25%", "21%", "", "21%", "", "", "20%", "20%", ""))
#reorder 
urb_rur_diff_2$short_name <- factor(urb_rur_diff_2$short_name,
                                    levels = c(unique(paste0(urb_rur_diff_2$short_name))))
levels(urb_rur_diff_2$short_name)
#side-side bar chart
urb_rur_viz_2 <- urb_rur_diff_2 %>% 
  ggplot(aes(reorder(short_name, urb_rur_diff), pct, fill = locale)) +
  geom_bar(stat = "identity", position = "dodge") +
  scale_fill_manual(values = paste(transcend_cols)) +
  theme_transcend_sparse +
  geom_text(aes(x = short_name, y = urb_rur_labs_2$y, label = urb_rur_labs_2$lab)) +
  theme(axis.text.x = element_text(angle = -55, hjust = 0)) +
  scale_y_continuous(limits = c(0, 100)) +
  labs(x = "",
       y = "Percent of Schools Selecting Tag",
       title = "Top 10 Tags with Largest Differences between Urban and Rural Schools",
       subtitle = "Comparison within locale types - denominator = schools within locale")

urb_rur_viz_2
```

### Urban vs. Suburban (Within)

```{r urb v sub viz 2, fig.dim = c(9, 9)}
urb_sub_diff_2 <- tags %>% 
  janitor::clean_names() %>% 
  select(variable_name, short_name) %>% 
  rename(tag = variable_name) %>% 
  left_join(locale_diffs_2, by = "tag") %>% 
  select(short_name, urb_sub_diff, pct_locale_tag_urban, pct_locale_tag_suburban) %>% 
  slice_max(urb_sub_diff, n = 10, with_ties = FALSE) %>% 
  mutate(urb_sub_diff = paste(round(1*urb_sub_diff, 0), "%"),
         pct_locale_tag_urban = 100*pct_locale_tag_urban,
         pct_locale_tag_suburban = 100*pct_locale_tag_suburban) %>% 
  pivot_longer(cols = c("pct_locale_tag_urban", "pct_locale_tag_suburban"),
               names_to = "locale",
               names_prefix = "pct_locale_tag_",
               values_to = "pct")
#pct labels
urb_sub_labs_2 <- data.frame(x = c(urb_sub_diff_2$short_name),
                             y = c(5 + urb_sub_diff_2$pct),
                             lab = c("", "31%", "", "28%", "", "26%", "", "25%", "23%", "", "", "20%", "", "20%", "19%", "", "19%", "", "", "18%"))
#reorder tags
urb_sub_diff_2$short_name <- factor(urb_sub_diff_2$short_name,
                                    levels = c(unique(paste0(urb_sub_diff_2$short_name))))
#side-side bar chart
urb_sub_viz_2 <- urb_sub_diff_2 %>% 
  ggplot(aes(short_name, pct, fill = locale)) +
  geom_bar(stat = "identity", position = "dodge") +
  ylim(0, 100) +
  geom_text(aes(x = short_name, y = urb_sub_labs_2$y, label = urb_sub_labs_2$lab)) +
  scale_fill_manual(values = paste(transcend_cols)) +
  theme_transcend_sparse +
  theme(axis.text.x = element_text(angle = -55, hjust = 0)) +
  labs(x = "",
       y = "Percent of Schools Selecting Tag",
       title = "Top 10 Tags with Largest Differences between Urban and Suburban Schools",
       subtitle = "Comparison within locale types - denominator = schools within locale")

urb_sub_viz_2
```

### Suburban vs. Rural (Within)  

```{r sub v rur viz 2, fig.dim = c(9, 9)}
sub_rur_diff_2 <- tags %>% 
  janitor::clean_names() %>% 
  select(variable_name, short_name) %>% 
  rename(tag = variable_name) %>% 
  left_join(locale_diffs_2, by = "tag") %>% 
  select(short_name, sub_rur_diff, pct_locale_tag_rural, pct_locale_tag_suburban) %>% 
  slice_max(sub_rur_diff, n = 10, with_ties = FALSE) %>% 
  mutate(sub_rur_diff = paste(round(1*sub_rur_diff, 0), "%"),
         pct_locale_tag_rural = 100*pct_locale_tag_rural,
         pct_locale_tag_suburban = 100*pct_locale_tag_suburban) %>% 
  pivot_longer(cols = c("pct_locale_tag_rural", "pct_locale_tag_suburban"),
               names_to = "locale",
               names_prefix = "pct_locale_tag_",
               values_to = "pct")
#pct labels
sub_rur_labs_2 <- data.frame(x = c(sub_rur_diff_2$short_name),
                           y = c(5 + sub_rur_diff_2$pct),
                           lab = c("", "43%", "", "33%", "", "32%", "", "32%", "", "29%", "", "29%", "", "29%", "", "25%", "", "25%", "", "25%"))
#reorder tags
sub_rur_diff_2$short_name <- factor(sub_rur_diff_2$short_name,
                                    levels = c(unique(paste0(sub_rur_diff_2$short_name))))
#side-side bar chart
sub_rur_viz_2 <- sub_rur_diff_2 %>% 
  ggplot(aes(short_name, pct, fill = locale)) +
  geom_bar(stat = "identity", position = "dodge") +
  ylim(0, 100) +
  geom_text(aes(x = short_name, y = sub_rur_labs_2$y, label = sub_rur_labs_2$lab)) +
  scale_fill_manual(values = paste(transcend_cols)) +
  theme_transcend_sparse +
  theme(axis.text.x = element_text(angle = -55, hjust = 0)) +
  labs(x = "",
       y = "Percent of Schools Selecting Tag",
       title = "Top 10 Tags with Largest Differences between Rural and Suburban Schools",
       subtitle = "Comparison within locale types - denominator = schools within locale")

sub_rur_viz_2
```

## Probability and Odds Ratios

In this final analysis, I calculated probability for each practice and converted to odds ratio in order to plot how much more or less likely a given locale was to select each tag.

```{r locale odds}
locale_odds <- locale_tags %>% 
  select(locale, tag, pct_school) %>% 
  pivot_wider(names_from = "locale",
              values_from = "pct_school",
              names_glue = "{.value}_{locale}") %>% 
  janitor::clean_names() %>% 
  mutate(odds_urban = pct_school_urban / (1 - pct_school_urban),
         odds_suburban = pct_school_suburban / (1 - pct_school_suburban),
         odds_rural = pct_school_rural / (1 - pct_school_rural),
         urb_rur_ratio = odds_urban / odds_rural,
         urb_sub_ratio = odds_urban / odds_suburban,
         sub_rur_ratio = odds_suburban / odds_rural)
```

### Urban vs. Rural (Odd Ratio)

```{r urb v rur viz 3, fig.dim = c(9, 9)}
urb_rur_odds <- tags %>% 
  janitor::clean_names() %>% 
  select(variable_name, short_name) %>% 
  rename(tag = variable_name) %>% 
  left_join(locale_odds, by = "tag") %>% 
  select(short_name, urb_rur_ratio) %>% 
  slice_max(urb_rur_ratio, n = 15, with_ties = FALSE)
#odd ratio labels labels
urb_rur_odds_lab <- data.frame(x = c(urb_rur_odds$short_name),
                           y = c(urb_rur_odds$urb_rur_ratio + .75),
                           lab = c(paste0(round(urb_rur_odds$urb_rur_ratio, 1), "x")))
#viz
urb_rur_odds_viz <- urb_rur_odds %>% 
  ggplot(aes(reorder(short_name, -urb_rur_ratio), urb_rur_ratio)) +
  geom_bar(stat = "identity", position = "dodge", fill = "#1A4C81") +
  coord_flip() +
  geom_text(aes(x = short_name, y = urb_rur_odds_lab$y, label = urb_rur_odds_lab$lab), 
            color = "#EF464B", size = 7, family = "Bebas Neue") +
  scale_color_manual(values = paste(transcend_cols)) +
  theme_transcend +
  theme(axis.text.x = element_blank(),
        axis.ticks.x = element_blank()) +
  labs(x = "",
       y = "",
       title = "How Much More Likely Urban Schools were to \n Select a Tag",
       subtitle = "as compared to rural schools using raw odds ratios")
  
urb_rur_odds_viz
```

### Urban vs. Suburban (Odd Ratio)

```{r urb v sub viz 3, fig.dim = c(9, 9)}
urb_sub_odds <- tags %>% 
  janitor::clean_names() %>% 
  select(variable_name, short_name) %>% 
  rename(tag = variable_name) %>% 
  left_join(locale_odds, by = "tag") %>% 
  select(short_name, urb_sub_ratio) %>% 
  slice_max(urb_sub_ratio, n = 16, with_ties = FALSE) %>% 
  filter(!c(short_name == "dual language programming"))
#odd ratio labels labels
urb_sub_odds_lab <- data.frame(x = c(urb_sub_odds$short_name),
                           y = c(.45 + urb_sub_odds$urb_sub_ratio),
                           lab = c(paste0(round(urb_sub_odds$urb_sub_ratio, 1), "x")))
#viz
urb_sub_odds_viz <- urb_sub_odds %>% 
  ggplot(aes(reorder(short_name, -urb_sub_ratio), urb_sub_ratio)) +
  geom_bar(stat = "identity", position = "dodge", fill = "#1A4C81") +
  coord_flip() +
  geom_text(aes(x = short_name, y = urb_sub_odds_lab$y, label = urb_sub_odds_lab$lab),
            color = "#EF464B", size = 7, family = "Bebas Neue") +
  scale_fill_manual(values = paste(transcend_cols)) +
  theme_transcend +
  theme(axis.text.x = element_blank(),
        axis.ticks.x = element_blank()) +
  labs(x = "",
       y = "",
       title = "How Much More Likely Urban Schools were to \n Select a Tag",
       subtitle = "as compared to suburban schools using raw odds ratios \n \n Note: dual language programming left out because no \n suburban schools selected it")
  
urb_sub_odds_viz
```

### Suburban vs. Rural (Odd Ratio)

```{r sub v rur viz 3, fig.dim = c(9, 9)}
sub_rur_odds <- tags %>% 
  janitor::clean_names() %>% 
  select(variable_name, short_name) %>% 
  rename(tag = variable_name) %>% 
  left_join(locale_odds, by = "tag") %>% 
  select(short_name, sub_rur_ratio) %>% 
  slice_max(sub_rur_ratio, n = 15, with_ties = FALSE)
#odd ratio labels labels
sub_rur_odds_lab <- data.frame(x = c(sub_rur_odds$short_name),
                           y = c(.15 + sub_rur_odds$sub_rur_ratio),
                           lab = c(paste0(round(sub_rur_odds$sub_rur_ratio, 1), "x")))
#viz
sub_rur_odds_viz <- sub_rur_odds %>% 
  ggplot(aes(reorder(short_name, -sub_rur_ratio), sub_rur_ratio)) +
  geom_bar(stat = "identity", position = "dodge", fill = "#1A4C81") +
  coord_flip() +
  geom_text(aes(x = short_name, y = sub_rur_odds_lab$y, label = sub_rur_odds_lab$lab),
            color = "#EF464B", size = 7, family = "Bebas Neue") +
  scale_fill_manual(values = paste(transcend_cols)) +
  theme_transcend +
  theme(axis.text.x = element_blank(),
        axis.ticks.x = element_blank()) +
  labs(x = "",
       y = "",
       title = "How Much More Likely Suburban Schools were to \n Select a Tag",
       subtitle = "as compared to rural schools using raw odds ratios")
  
sub_rur_odds_viz
```

Save plots:
```{r, eval = FALSE}
#across plots
ggsave("urb-rur_across.png", plot = urb_rur_viz, path = here("outputs", "locale analysis"),
       width = 12, height = 8, units = "in")
ggsave("urb-sub_across.png", plot = urb_sub_viz, path = here("outputs", "locale analysis"),
       width = 12, height = 8, units = "in")
ggsave("sub-rur_across.png", plot = urb_sub_viz, path = here("outputs", "locale analysis"),
       width = 12, height = 8, units = "in")
#within plots
ggsave("urb-rur_within.png", plot = urb_rur_viz_2, path = here("outputs", "locale analysis"),
       width = 12, height = 8, units = "in")
ggsave("urb-sub_within.png", plot = urb_sub_viz_2, path = here("outputs", "locale analysis"),
       width = 12, height = 8, units = "in")
ggsave("sub-rur_within.png", plot = sub_rur_viz_2, path = here("outputs", "locale analysis"),
       width = 12, height = 8, units = "in")
#odds charts
ggsave("urb-rur_odds_chart_raw.png", plot = urb_rur_odds_viz, path = here("outputs", "locale analysis"),
       width = 12, height = 8, units = "in")
ggsave("urb-sub_odds_chart_raw.png", plot = urb_sub_odds_viz, path = here("outputs", "locale analysis"),
       width = 12, height = 8, units = "in")
ggsave("sub-rur_odds_chart_raw.png", plot = sub_rur_odds_viz, path = here("outputs", "locale analysis"),
       width = 12, height = 8, units = "in")
```

