---
title: "Do School Leaders Represent the Students they Serve?"
author: "Anwesha Guha"
date: "6/15/2022"
output: 
  html_document:
    toc: true
    toc_float: true
    theme: flatly
    highlight: tango
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Setup for analysis

```{r}
library(pacman)
p_load(rio, here, psych, tidyverse, DT, scales, extrafont)
source(here("scripts", "branding.R"))
load(here("data", "complete_canopy_2022.RData"))
```

## Overview

In this analysis, we are trying to answer the following question: *Do school leaders represent the students they serve?*

Let's look at a school-level.

Here are the demographic distributions for each school.

Only looking at those with less than 10% White students:

```{r}
demos <- demographic_data %>% 
  mutate(enrollment = case_when(
    nces_total_enrollment > 0 ~ nces_total_enrollment,
    TRUE ~ self_reported_total_enrollment)) %>% 
  mutate(bipoc_percent = 1 - white_percent) %>% 
  arrange(-bipoc_percent) %>% 
  filter(enrollment > 10) %>% 
  select(confidential_leader_race, confidential_leadership_team_diversity, confidential_teaching_staff_diversity, bipoc_percent, enrollment)

datatable(demos, colnames = c("School Leader Race", "Leadership Team", "Teaching Staff", "Pct. BIPOC Students", "Enrollment")) %>% 
  formatPercentage(c("bipoc_percent"), 2) 
```

Teaching Staff
```{r}
demos %>% 
  filter(confidential_teaching_staff_diversity != "0" 
         & confidential_teaching_staff_diversity != "Prefer not to say"
         & confidential_leader_race != "NA") %>% 
  ggplot(aes(confidential_teaching_staff_diversity, confidential_leader_race)) +
  geom_tile(aes(fill = bipoc_percent)) +
  scale_fill_gradient(low = "pink", high = "purple") +
  labs(x = "Teaching Staff",
       y = "School Leader Race") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5)) +
  coord_flip()
```

Leadership Team
```{r}
demos %>% 
  filter(confidential_leadership_team_diversity != "0" 
         & confidential_leadership_team_diversity != "Prefer not to say"
         & confidential_leadership_team_diversity != "Not sure"
         & confidential_leader_race != "NA") %>%
  ggplot(aes(confidential_leadership_team_diversity, confidential_leader_race)) +
  geom_tile(aes(fill = bipoc_percent)) +
  labs(x = "Leadership Team",
       y = "School Leader Race") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5)) +
  scale_fill_continuous(low = "pink", high = "purple", na.value = "gray") +
  coord_flip()
```

```{r}
table(demographic_data$confidential_leader_race)
```


```{r}
table(demographic_data$confidential_leadership_team_diversity)
```

```{r}
table(demographic_data$confidential_teaching_staff_diversity)
```

**Notes:**
do 1- white pct 
number of schools falling into categories (and then dot for size of school)
filter out small schools
what pct of students have same race/ethnicity 

bipoc leadership vs bipoc student count
focus on hispanic and black

look at everyone above median for black stuednts and look at how many have black leaders
y - same leader race pct, x - bins of students 

# Representation Visualizations v2

## Black students

```{r}
demo2 <- demographic_data %>% 
  mutate(enrollment = case_when(
    nces_total_enrollment > 0 ~ nces_total_enrollment,
    TRUE ~ self_reported_total_enrollment)) %>% 
  mutate(bipoc_percent = 1 - white_percent) %>% 
  arrange(-bipoc_percent) %>% 
  filter(enrollment > 10) %>% 
  select(confidential_leader_race, confidential_leadership_team_diversity, confidential_teaching_staff_diversity, bipoc_percent, enrollment, black_percent, hispanic_percent)
```

```{r}
hist(demo2$black_percent)
```

```{r}
hist(demo2$hispanic_percent)
```


```{r black demo viz, fig.height=8, fig.width=8}
demo2 %>% 
  filter(black_percent > .5 & 
           confidential_leadership_team_diversity != "Prefer not to say") %>% 
  rename(`school leader race` = confidential_leader_race) %>% 
  ggplot(aes(x = confidential_leadership_team_diversity, y = confidential_teaching_staff_diversity)) +
  geom_jitter(aes(size = enrollment, color = `school leader race`), alpha = .6) +
  theme(legend.position = c(0.8, 0.2),
        legend.text = element_text(size = 10),
        legend.title = element_text(size = 10),
        legend.key = element_rect(fill = "transparent"),
        axis.text = element_text(size = 10),
        axis.text.x = element_text(angle = 40, vjust = 0.5),
        plot.caption = element_text(size = 10, hjust = 0))  +
  labs( x = "Leadership Team",
        y = "Teaching Staff",
        title = "For schools with more than 50% Black \nstudents, what do their leaders look like?",
        caption = "Of the 16 schools included, there were 10 Black, 5 White, 3 Black, and 1 Biracial \nschool leaders. Size of points determined by school enrollment.") +
  guides(size = "none")
```

```{r hispanic demo viz, fig.height=8, fig.width=8}
demo2 %>% 
  filter(hispanic_percent > .5 & 
           confidential_leader_race != "No Response" & 
           confidential_leadership_team_diversity != "Not sure") %>% 
  rename(`school leader race` = confidential_leader_race) %>% 
  ggplot(aes(x = confidential_leadership_team_diversity, y = confidential_teaching_staff_diversity)) +
  geom_jitter(aes(size = enrollment, color = `school leader race`), alpha = .6) +
  theme(legend.position = c(0.8, 0.2),
        legend.text = element_text(size = 10),
        legend.title = element_text(size = 10),
        legend.key = element_rect(fill = "transparent"),
        axis.text = element_text(size = 10),
        axis.text.x = element_text(angle = 40, vjust = 0.5),
        plot.caption = element_text(size = 10, hjust = 0))  +
  labs( x = "Leadership Team",
        y = "Teaching Staff",
        title = "For schools with more than 50% Hispanic \nstudents, what do their leaders look like?",
        caption = "Of the 30 schools included, there were 17 White, 8 Hispanic, 3 Black, and 1 Asian \nand Xicana Indigena school leaders. Size of points determined by school enrollment.") +
  guides(size = "none")
```




